### 2018-11-06 gg_topi.Rmd
Some changes

---
title: "GG_topi2"
author: "Francesco Beghini"
date: "26/09/2018"
output: 
  html_document: 
    theme: cerulean
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(phyloseq)
library(magrittr)
library(tidyverse)
library(reshape2)
library(ggsignif)
library(DESeq2)

```

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/otu_table_mc10_w_tax.biom',
                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/rep_set.tre',
                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/rep_set.fna',
                        refseqFunction = parse_taxonomy_default
                      )

mappingfile <- read.delim('/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/mappingfile', row.names = 1, check.names = FALSE)
mappingfile$Description %<>% str_replace_all(.,'_','-')
mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
rownames(mappingfile) <- mappingfile$Description

sample_data(gg_topi) <- mappingfile
sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)
sample_data(gg_topi)$pentatome <- factor(stringr::str_sub(sample_data(gg_topi)$condizione,1,2), labels = c('Pentatoma','Empty'))
sample_data(gg_topi)$bifido <- factor(stringr::str_sub(sample_data(gg_topi)$condizione,3,3), labels = c('Bifido','PBS'))
sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))

colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")

tax_table_filtered <- tax_table(gg_topi) %>%
  as.data.frame %>% 
  mutate_all( .funs = function(x) stringr:::str_replace_all(x,"D_.__",""))

rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
tax_table(gg_topi) <-  as.matrix(tax_table_filtered)

gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

gg_topi.phylum <- tax_glom(gg_topi,taxrank = "Phylum")
gg_topi.phylum.relab <- transform_sample_counts(gg_topi.phylum, function(x) x/sum(x))

meta_samples <- data.frame(otu_table(gg_topi), check.names = FALSE) %>% 
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  tibble::add_column(., timepoint = stringr::str_sub(.$SampleID, 12, 13)) %>%
  dplyr::group_by(timepoint, OTU) %>%
  dplyr::select(-SampleID)%>%
  dplyr::summarise(Abundance=mean(Abundance)) %>%
  tidyr::spread(timepoint, Abundance) %>%
  tibble::column_to_rownames(var = 'OTU') %>%
  data.frame()

gg_topi_meta_samples <- phyloseq(otu_table(meta_samples, taxa_are_rows = TRUE), tax_table(gg_topi), phy_tree(gg_topi))
```
<!-- # Taxonomy profile at genus level -->
<!-- ```{r taxonomy, cache=FALSE} -->
<!-- # nychanesmicrobiome::plot_abundance(gg_topi,taxrank = 'Genus') -->
<!-- ``` -->

VVP: Vuote + PBS

VPP: Pentatoma + PBS

VVB: Vuote + *Bifidobacterium*

VPB: Pentatoma + *Bifidobacterium*

# Alpha diversity
```{r Alpha diversity}
# otu_table(gg_topi) %>% colSums() %>% sort %>% barplot(xlab = c())
# otu_table(gg_topi) %>% colSums() %>% summary()

alpha_div <- estimate_richness(gg_topi)
alpha_div %<>% bind_cols(data.frame(sample_data(gg_topi)))

x <- melt(alpha_div, id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon'))
x$value <- as.numeric(x$value)

ggplot(x) +
  geom_boxplot(aes(timepoint, value)) +
  facet_grid(variable~condizione,scales = 'free_y')
```
<!-- ## SFB abundance -->
<!-- ```{r} -->
<!-- knitr::kable(otu_table(gg_topi)[c(879,880),]) -->
<!-- ``` -->

# Beta diversity
```{r Beta diversity, cache=TRUE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
distbc <- phyloseq::distance(gg_topi, "bray")
distjac <- phyloseq::distance(gg_topi, "jaccard")
distwu_meta <- phyloseq::distance(gg_topi_meta_samples, "jaccard")

ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordwu_meta <-  ordinate(gg_topi_meta_samples, method = "MDS", distance = distwu_meta, weighted = TRUE)
ordbc <-  ordinate(gg_topi, method = "MDS", distance = distbc, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordjac <-  ordinate(gg_topi, method = "MDS", distance = distjac, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

# plot_ordination(gg_topi, ordwu, color = 'condizione',shape = 'timepoint', title = 'Weighted UniFrac')
# plot_ordination(gg_topi, ordwu, color = 'merge_timepoint', title = 'Weighted UniFrac')

distances <- reshape2::melt(as.matrix(distwu), varnames = c('rows','cols'))
distances$condizione1 <- sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][2])
distances$pentatome1 <- sapply(as.character(distances$rows), function(x) str_sub(x, 5,5))
distances$condizione2 <- sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][2])
distances$topo1 <- factor(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][3]))
distances$topo2 <- factor(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][3]))
distances$timepoint1 <- stringr::str_replace(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distances$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric

distancesbc <- reshape2::melt(as.matrix(distbc), varnames = c('rows','cols'))
distancesbc$condizione1 <- sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][2])
distancesbc$condizione2 <- sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][2])
distancesbc$topo1 <- factor(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][3]))
distancesbc$topo2 <- factor(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][3]))
distancesbc$timepoint1 <- stringr::str_replace(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distancesbc$timepoint2 <- stringr::str_replace(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric

distancesjac <- reshape2::melt(as.matrix(distjac), varnames = c('rows','cols'))
distancesjac$condizione1 <- sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][2])
distancesjac$condizione2 <- sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][2])
distancesjac$topo1 <- factor(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$topo2 <- factor(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][3]))
distancesjac$timepoint1 <- stringr::str_replace(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
```

## Ordination: Mean all samples per timepoint
```{r}

x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)

ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
  geom_point(size = 0.8) +
  geom_point(data=centroids, size = 4)
```

## Ordination: T2 vs T3, tumor vs no tumor
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T2','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T2','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=tumor_injected), size=3, alpha=.8) +
  scale_color_brewer(name = "Tumor injected", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = -0.2, xmax = -.2, ymin = .25, ymax = .25)
```

## Alpha diversity: tumor vs no tumor
```{r}
filt_alpha <- alpha_div %>% 
  filter(timepoint %in% c('T2','T3')) %>%
  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2)))

ttest_res <- data.frame()
for(measure in c('Observed', 'se.chao1', 'Simpson', 'Shannon')){
  ttest_res %<>% bind_rows(data.frame(Measure=measure, pvalue=with(filt_alpha, t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
}

filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon')) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, value)) +
  facet_grid(variable~.,scales = 'free_y')

knitr::kable(ttest_res)
```

### Merge T1 and T2 and use them as baseline
## Ordination: baseline vs T3
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T3'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T4
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T4'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T5
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T5'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T6
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T6'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T7
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T7'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: T3 vs T7: Treated with Bifidum?
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('T3','T7'))) +
  geom_point(aes(Axis.1, Axis.2, color=cond_all), size=3) +
  scale_colour_discrete(name="Timepoint_Vaccine_Bifido") +
  theme_bw() 
```

## Ordination: T3 vs T4
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T4','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T4','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ timepoint + real_vaccine + timepoint:real_vaccine, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T4','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint, shape=real_vaccine), size=3, alpha=.7) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  scale_shape_manual(name = 'Pentatome?', values=c(15,19))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('Timepoint PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = .2, xmax = .2, ymin = .1, ymax = .1) +
  annotation_custom(grob = grid::textGrob(paste('Pentatome PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[2], '\nR2=', signif(permanova$aov.tab$R2[2], 2)), 
gp=grid::gpar(col="black", fontsize=10)),  xmin = -.1, xmax = -.1, ymin = -.08, ymax = -.08)
```

## Ordination: T3 vs T5
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T5','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T5','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T5','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .12, ymax = .12)
```

## Ordination: T3 vs T6
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T6','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T6','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T6','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
```

## Ordination: T3 vs T7
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
```

## Beta diversity between timepoints, T1 is the baseline
```{r Beta diversity between timepoints t1}
distances[distances$timepoint1 == 1 | distances$timepoint1 == 2, 'timepoint1'] <- 2
distances[distances$timepoint2 == 1 | distances$timepoint2 == 2, 'timepoint2'] <- 2

t1 <- distances %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==2) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

t.res <- data.frame()
for(cond in unique(t1$condizione1)){
  for(x in combn(c(3:7), 2, simplify = FALSE)){
    pval.t = t.test(t1[t1$timepoint2==x[1] & t1$condizione1==cond,"value"], t1[t1$timepoint2==x[2] & t1$condizione1==cond ,"value"])$p.value
    pval.w = wilcox.test(t1[t1$timepoint2==x[1] & t1$condizione1==cond,"value"], t1[t1$timepoint2==x[2] & t1$condizione1==cond ,"value"], paired = F)$p.value
    if(pval.t < 0.05 | pval.w < 0.05)
      t.res %<>% bind_rows(c(condizione1=cond, timepoint=as.numeric(x), p.value = str_c('t-test:',signif(pval.t, digits = 2),'U-test:', signif(pval.w, digits = 2), sep = ' ')))
    }
  }

```

```{r}
knitr::kable(t.res)
```

```{r}

t1 %>%
  ggplot(aes(timepoint2, value, group = timepoint2)) +
  stat_boxplot() +
  scale_x_continuous(breaks = c(2:7)) +
  # geom_signif(data=t.res,
              # aes(xmin= as.numeric(timepoint1), xmax = as.numeric(timepoint2), annotations = p.value, y_position = 0.33),
              # manual = TRUE) +
  facet_grid(condizione1 ~ .)

```

## Beta diversity between timepoints, T(-1) is the baseline
```{r Beta diversity between timepoints tn-1}
tn1 <- distances %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==timepoint2-1) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

t.res <- data.frame()
for(cond in unique(tn1$condizione1)){
  for(x in combn(c(3:7), 2, simplify = FALSE)){
      pval.t = t.test(tn1[tn1$timepoint2==x[1] & tn1$condizione1==cond,"value"], tn1[tn1$timepoint2==x[2] & tn1$condizione1==cond ,"value"])$p.value
      pval.w = wilcox.test(tn1[tn1$timepoint2==x[1] & tn1$condizione1==cond,"value"], tn1[tn1$timepoint2==x[2] & tn1$condizione1==cond ,"value"], paired = F)$p.value
      if(pval.t < 0.05 | pval.w < 0.05)
        t.res %<>% bind_rows(c(condizione1=cond, timepoint=as.numeric(x), p.value = str_c('t-test:',signif(pval.t, digits = 2),'U-test:', signif(pval.w, digits = 2), sep = ' ')))
  }
}
```

```{r}
knitr::kable(t.res)
```

```{r}
tn1 %>%
  ggplot(aes(timepoint2, value, group = timepoint2)) +
  geom_boxplot() +
  scale_x_continuous(breaks = c(2:7)) +
  # geom_signif(data=t.res,
              # aes(xmin= as.numeric(timepoint1), xmax = as.numeric(timepoint2), annotations = p.value, y_position = 0.33),
              # manual = TRUE) +
  facet_grid(condizione1 ~ .)
```

## Differential analysis
```{r DA, cache=TRUE}
threshold <- 0.05

ggtopi_filt <- filter_taxa(gg_topi, function(x) mean(x) > 5e-1, TRUE)

dds <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ cond_vac), parallel = TRUE, quiet = TRUE, fitType = 'parametric')

dds_bif <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi, design = ~ cond_bif), parallel = TRUE, quiet = TRUE)
dds1 <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ timepoint), parallel = TRUE, quiet = TRUE)

dds_phylum <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi.phylum, design = ~ timepoint), parallel = TRUE, quiet = TRUE)

```
## T2 vs T3
```{r DA T2 vs T3}
res23 <- DESeq2::results(dds1, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
res23.filtered <- res23[!is.na(res23$padj),]
res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]
res23.filtered <- res23.filtered[abs(res23.filtered$log2FoldChange) > 2,]

res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = log(baseMean))) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('T1 vs T2') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

## T2 vs T3 phylum level
```{r DA T2 vs T3 phylum}
res_phylum <- DESeq2::results(dds_phylum, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
res_phylum.filtered <- res_phylum[!is.na(res_phylum$padj),]
res_phylum.filtered <- res_phylum.filtered[res_phylum.filtered$padj < threshold,]
# res_phylum.filtered <- res_phylum.filtered[abs(res_phylum.filtered$log2FoldChange) > 2,]

res_phylum.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res_phylum.filtered),])) %>%
  ggplot(aes(log2FoldChange, Phylum)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

## Check T1 vs T2
```{r DA T1 vs T2}
res12 <- DESeq2::results(dds1, contrast = c("timepoint","T1","T2"))
res12.filtered <- res12[!is.na(res12$padj),]
res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

res12.filtered %>% 
  as.data.frame %>% 
  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('Timepoint 1 vs Timepoint 2') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

```{r DA T3 vs T4}
res23 <- DESeq2::results(dds1, contrast = c("timepoint","T3","T4"))
res23.filtered <- res23[!is.na(res23$padj),]
res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]

res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
  ggplot(aes(log2FoldChange, Genus)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  ggtitle('Timepoint 3 vs Timepoint 4') +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
)
```

## Differential analysis Pentatome T7 vs not vaccinated T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"), independentFiltering = TRUE)
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = log10(baseMean))) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Pentatome T3 vs T7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

```

## Differential analysis Pentatome T3 VS T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = log10(baseMean))) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Pentatome T3 vs T7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

```

```{r}
res12.filtered %>% 
  as.data.frame %>% 
  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
  filter(is.na(Phylum)) %>%  knitr::kable()
```

## Differential analysis Empty vescicles T3 VS T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_NO","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle('Empty T3vsT7') +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

```

## Differential analysis Pentatoma vs empty vescicles T7
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

```

## Differential analysis Pentatoma vs empty vescicles T3
```{r}
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T3_YES","T3_NO"))
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
  timepoint_number <- stringr::str_sub(timepoint,17,17)

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))

```

## Differential analysis Pentatome vs No-pentatome at different timepoints
```{r}
for (timepoint in c("merge_timepointT3.pentatomePBS","merge_timepointT4.pentatomePBS","merge_timepointT5.pentatomePBS","merge_timepointT6.pentatomePBS","merge_timepointT7.pentatomePBS")) {
  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T6_Pentatoma","T6_Empty"), )
  res12.filtered <- res12[!is.na(res12$padj),]
  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
  timepoint_number <- stringr::str_sub(timepoint,17,17)

  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
    ggplot(aes(log2FoldChange, Genus)) +
    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
    geom_point(size=3, aes(color = baseMean)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
    theme_bw() +
    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
    scale_y_discrete(position = "right") +
    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
          strip.text = element_text(size = 9),
          axis.title = element_text(size = 9),
          axis.title.y = element_blank(),
          strip.background = element_rect(colour = "transparent", fill = "transparent"),
          strip.text.y = element_text(angle = 180)
    ))
}
```

<!-- ## Differential analysis Bifido vs No-bifido at different timepoints -->
<!-- ```{r} -->
<!-- for (timepoint in c("bifidoPBS.merge_timepointT3","bifidoPBS.merge_timepointT4","bifidoPBS.merge_timepointT5","bifidoPBS.merge_timepointT6","bifidoPBS.merge_timepointT7")) { -->
<!--   res12 <- DESeq2::results(dds2, contrast = list(c("bifido_PBS_vs_Bifido" , timepoint))) -->
<!--   res12.filtered <- res12[!is.na(res12$padj),] -->
<!--   res12.filtered <- res12.filtered[res12.filtered$padj < threshold,] -->
<!--   timepoint_number <- stringr::str_sub(timepoint,-1) -->

<!--   plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>% -->
<!--     ggplot(aes(log2FoldChange, Genus)) + -->
<!--     geom_vline(xintercept = 0.0, color = "gray", size = 0.5) + -->
<!--     geom_point(size=3, aes(color = baseMean)) + -->
<!--     facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+ -->
<!--     theme_bw() + -->
<!--     ggtitle(paste('Bifido vs no bifido at timepoint', timepoint_number, sep = ' ')) + -->
<!--     scale_y_discrete(position = "right") + -->
<!--     theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11), -->
<!--           strip.text = element_text(size = 9), -->
<!--           axis.title = element_text(size = 9), -->
<!--           axis.title.y = element_blank(), -->
<!--           strip.background = element_rect(colour = "transparent", fill = "transparent"), -->
<!--           strip.text.y = element_text(angle = 180) -->
<!--     )) -->
<!-- } -->
<!-- ``` -->

# PERMANOVA
```{r}
metadata <- as(sample_data(gg_topi), "data.frame")
permanova.res <- data.frame()
distwu <- phyloseq::distance(gg_topi, "wunifrac")

all_permanova <- as.data.frame(vegan::adonis(formula = distwu ~ merge_timepoint*bifido*pentatome, data = metadata)$aov.tab)
```

```{r}
knitr::kable(all_permanova)
```

```{r}
for(s in combn(levels(metadata$merge_timepoint),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$merge_timepoint %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r}
for(s in combn(levels(metadata$pentatome),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$pentatome %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, pentatome %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ pentatome, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r}
for(s in combn(levels(metadata$bifido),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$bifido %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, bifido %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ bifido, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r}
knitr::kable(permanova.res)
```

## PERMANOVA all vs all
```{r}
permanova.res <- data.frame()
for (s in combn(levels(sample_data(gg_topi)$cond_all),2, simplify = FALSE)) {
  metadata_2cat <- data.frame(sample_data(gg_topi))[sample_data(gg_topi)$cond_all %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, cond_all %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r}
knitr::kable(permanova.res)
```

## (T5+T6+T7) vs baseline
```{r}
metadata_tub <- metadata[metadata$merge_timepoint %in% c('baseline','T5','T6','T7'),]
metadata_tub$merge_timepoint <- as.character(metadata_tub$merge_timepoint)
metadata_tub[which(metadata_tub$merge_timepoint!='baseline'), 'merge_timepoint'] <- 'Timm'
distwu_tub <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% c('baseline','T5','T6','T7')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_tub ~ merge_timepoint, data = metadata_tub)
```

```{r}
knitr::kable(permanova$aov.tab)
```

## Effect of bifido in T4 vs T5
```{r}
metadata_t45 <- metadata[metadata$merge_timepoint %in% c('T4','T5'),]
distwu_t45 <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% c('T4','T5')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_t45 ~ merge_timepoint*bifido, data = metadata_t45)
```

```{r}
knitr::kable(permanova$aov.tab)
```

## Beta diversity T3 vs T4
```{r}
t1 <- distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2==4) %>%
  dplyr::arrange(topo1)

t1 %>%
  ggplot(aes(pentatome1, value)) +
  stat_boxplot() +
  stat_signif()
```

## Beta diversity T3 vs T7
```{r}
t37 <- distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2==7) %>%
  dplyr::arrange(topo1)

t37 %>%
  ggplot(aes(condizione1, value)) +
  stat_boxplot()
```

## Correlation beta-diversity T2-TX, tumor mass
```{r corr tumor wunifrac}
distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
wunif_ds <- distances %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

ds_corr_wunif <- data.frame()
for(x in unique(wunif_ds$timepoint2)){
  for(y in levels(wunif_ds$condizione)){
    tmp_res <- with(wunif_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
    ds_corr_wunif %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
  }
}

knitr::kable(ds_corr_wunif)
```

```{r corr tumor braycurtis}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
brayc_ds <- distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

ds_corr_braycf <- data.frame()
for(x in unique(brayc_ds$timepoint2)){
  for(y in levels(brayc_ds$condizione)){
    tmp_res <- with(brayc_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
    ds_corr_braycf %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
  }
}

knitr::kable(ds_corr_braycf)
```

```{r corr tumor jaccard}
distancesjac %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
jac_ds <- distancesjac %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==2) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

ds_corr_jac <- data.frame()
for(x in unique(jac_ds$timepoint2)){
  for(y in levels(jac_ds$condizione)){
    tmp_res <- with(jac_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
    ds_corr_jac %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
  }
}

knitr::kable(ds_corr_jac)
```

# Tracking OTU
```{r}
counts <- data.frame(as.data.frame(otu_table(gg_topi)[c('EU622777.1.1487','EU504401.1.1405'),] %>% t)) %>% tibble::rownames_to_column('sampleid')
colnames(counts) <- c('sampleid',"Lachnospiraceae_NK4A136","Muribaculaceae")
counts$t <- sapply(as.character(counts$sampleid), function(x) strsplit(x,'-')[[1]][4])
counts$condizione <- sapply(as.character(counts$sampleid), function(x) strsplit(x,'-')[[1]][2])

reshape2::melt(counts, id.vars = c('t', 'condizione','sampleid')) %>%
  ggplot() +
  stat_boxplot(aes(t, value)) + 
  facet_grid(variable ~ condizione, scales = "free_x", space="free_y")
```

### 2018-11-07 gg_topi.Rmd
A

@@
 sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)
-sample_data(gg_topi)$pentatome <- factor(stringr::str_sub(sample_data(gg_topi)$condizione,1,2), labels = c('Pentatoma','Empty'))
-sample_data(gg_topi)$bifido <- factor(stringr::str_sub(sample_data(gg_topi)$condizione,3,3), labels = c('Bifido','PBS'))
+
+sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
+sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])
+
 sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
@@
 sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))
+
+sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)
 
@@
 ```{r}
-
 x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
@@
   scale_colour_discrete(name="Timepoint_Vaccine_Bifido") +
+  theme_bw() 
+```
+
+```{r}
+ggplot(subset(ordwu, timepoint %in% c('T3','T7'))) +
+  geom_point(aes(Axis.1, Axis.2, color=timepoint, shape=bifido), size=3) +
+  facet_wrap(pentatome~bifido)+
   theme_bw() 
@@
 
+mmtx <- model.matrix(object = ~timepoint*real_vaccine*real_bifido, data = data.frame(gg_topi@sam_data))
+idx <- which(apply(mmtx, 2, function(x) all(x==0)))
+mmtx <- mmtx[,-idx]
+
+rmmtx <- model.matrix(object = ~ timepoint, data = data.frame(gg_topi@sam_data))
+
+pds <- phyloseq_to_deseq2(ggtopi_filt, design = ~timepoint*real_vaccine*real_bifido, ignoreRank = TRUE)
+
+dds <- DESeq2::DESeq(pds, full = mmtx, reduced = rmmtx, test = 'LRT')
+
 dds_bif <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi, design = ~ cond_bif), parallel = TRUE, quiet = TRUE)
@@
 ```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"), independentFiltering = TRUE)
+  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
+  res12 <- DESeq2::results(dds)
   res12.filtered <- res12[!is.na(res12$padj),]

### 2018-11-19 cage_effect.Rmd
T7 no 24/7

--- 2018-11-06 gg_topi.Rmd
@@
 ---
-title: "GG_topi2"
+title: "Cage effect"
 author: "Francesco Beghini"
-date: "26/09/2018"
-output: 
-  html_document: 
-    theme: cerulean
+date: "November 8, 2018"
+output: html_document
 editor_options: 
@@
 sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)
-sample_data(gg_topi)$pentatome <- factor(stringr::str_sub(sample_data(gg_topi)$condizione,1,2), labels = c('Pentatoma','Empty'))
-sample_data(gg_topi)$bifido <- factor(stringr::str_sub(sample_data(gg_topi)$condizione,3,3), labels = c('Bifido','PBS'))
+
+sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
+sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])
 sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
@@
 tax_table(gg_topi) <-  as.matrix(tax_table_filtered)
+```
 
-gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
-gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))
-
-gg_topi.phylum <- tax_glom(gg_topi,taxrank = "Phylum")
-gg_topi.phylum.relab <- transform_sample_counts(gg_topi.phylum, function(x) x/sum(x))
-
-meta_samples <- data.frame(otu_table(gg_topi), check.names = FALSE) %>% 
-  tibble::rownames_to_column("OTU") %>%
-  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
-  tibble::add_column(., timepoint = stringr::str_sub(.$SampleID, 12, 13)) %>%
-  dplyr::group_by(timepoint, OTU) %>%
-  dplyr::select(-SampleID)%>%
-  dplyr::summarise(Abundance=mean(Abundance)) %>%
-  tidyr::spread(timepoint, Abundance) %>%
-  tibble::column_to_rownames(var = 'OTU') %>%
-  data.frame()
-
-gg_topi_meta_samples <- phyloseq(otu_table(meta_samples, taxa_are_rows = TRUE), tax_table(gg_topi), phy_tree(gg_topi))
-```
-<!-- # Taxonomy profile at genus level -->
-<!-- ```{r taxonomy, cache=FALSE} -->
-<!-- # nychanesmicrobiome::plot_abundance(gg_topi,taxrank = 'Genus') -->
-<!-- ``` -->
-
-VVP: Vuote + PBS
-
-VPP: Pentatoma + PBS
-
-VVB: Vuote + *Bifidobacterium*
-
-VPB: Pentatoma + *Bifidobacterium*
-
-# Alpha diversity
-```{r Alpha diversity}
-# otu_table(gg_topi) %>% colSums() %>% sort %>% barplot(xlab = c())
-# otu_table(gg_topi) %>% colSums() %>% summary()
-
-alpha_div <- estimate_richness(gg_topi)
-alpha_div %<>% bind_cols(data.frame(sample_data(gg_topi)))
-
-x <- melt(alpha_div, id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon'))
-x$value <- as.numeric(x$value)
-
-ggplot(x) +
-  geom_boxplot(aes(timepoint, value)) +
-  facet_grid(variable~condizione,scales = 'free_y')
-```
-<!-- ## SFB abundance -->
-<!-- ```{r} -->
-<!-- knitr::kable(otu_table(gg_topi)[c(879,880),]) -->
-<!-- ``` -->
-
-# Beta diversity
+# How's the dispostion of cages
 ```{r Beta diversity, cache=TRUE}
 distwu <- phyloseq::distance(gg_topi, "wunifrac")
-distbc <- phyloseq::distance(gg_topi, "bray")
-distjac <- phyloseq::distance(gg_topi, "jaccard")
-distwu_meta <- phyloseq::distance(gg_topi_meta_samples, "jaccard")
+ordwu <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE)
+```
 
-ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
-ordwu_meta <-  ordinate(gg_topi_meta_samples, method = "MDS", distance = distwu_meta, weighted = TRUE)
-ordbc <-  ordinate(gg_topi, method = "MDS", distance = distbc, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
-ordjac <-  ordinate(gg_topi, method = "MDS", distance = distjac, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
+```{r}
+plot_ordination(gg_topi, ordwu, color="cage")
+```
 
-# plot_ordination(gg_topi, ordwu, color = 'condizione',shape = 'timepoint', title = 'Weighted UniFrac')
-# plot_ordination(gg_topi, ordwu, color = 'merge_timepoint', title = 'Weighted UniFrac')
-
+# B-diversity intra cage
+```{r, cache = TRUE}
 distances <- reshape2::melt(as.matrix(distwu), varnames = c('rows','cols'))
-distances$condizione1 <- sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][2])
-distances$pentatome1 <- sapply(as.character(distances$rows), function(x) str_sub(x, 5,5))
-distances$condizione2 <- sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][2])
 distances$topo1 <- factor(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][3]))
@@
 
-distancesbc <- reshape2::melt(as.matrix(distbc), varnames = c('rows','cols'))
-distancesbc$condizione1 <- sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][2])
-distancesbc$condizione2 <- sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][2])
-distancesbc$topo1 <- factor(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][3]))
-distancesbc$topo2 <- factor(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][3]))
-distancesbc$timepoint1 <- stringr::str_replace(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-distancesbc$timepoint2 <- stringr::str_replace(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-
-distancesjac <- reshape2::melt(as.matrix(distjac), varnames = c('rows','cols'))
-distancesjac$condizione1 <- sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][2])
-distancesjac$condizione2 <- sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][2])
-distancesjac$topo1 <- factor(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][3]))
-distancesjac$topo2 <- factor(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][3]))
-distancesjac$timepoint1 <- stringr::str_replace(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
+distances %<>%
+  inner_join(unique(data.frame(sample_data(gg_topi))[,c('id.topo','cage')]), by = c('topo1' = 'id.topo')) %>%
+  rename('cage'='cage1') %>%
+  left_join(unique(data.frame(sample_data(gg_topi))[,c('id.topo','cage')]), by = c('topo2' = 'id.topo')) %>%
+  rename('cage'='cage2')
 ```
 
-## Ordination: Mean all samples per timepoint
+# Intra cage diversity through time
 ```{r}
-
-x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
-
-centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)
-
-ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
-  geom_point(size = 0.8) +
-  geom_point(data=centroids, size = 4)
+  distances %>%
+  filter(cage1==cage2) %>%
+  filter(timepoint1==timepoint2) %>%
+  ggplot() +
+  geom_boxplot(aes(cage1, value)) + 
+  facet_wrap(timepoint1 ~ .)
 ```
 
-## Ordination: T2 vs T3, tumor vs no tumor
+# Inter cage diversity through time
 ```{r}
-metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T2','T3'),]
-distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T2','T3')), "wunifrac")
-permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
-
-subset(ordwu, timepoint %in% c('T2','T3')) %>%
-ggplot(aes(Axis.1, Axis.2)) +
-  geom_point(aes(color=tumor_injected), size=3, alpha=.8) +
-  scale_color_brewer(name = "Tumor injected", palette="Dark2") + 
-  theme_minimal() +
-  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = -0.2, xmax = -.2, ymin = .25, ymax = .25)
+  distances %>%
+  filter(cage1!=cage2) %>%
+  filter(timepoint1==timepoint2) %>%
+  mutate(cage = paste(.$cage1, .$cage2, sep =' vs ')) %>%
+  ggplot() +
+  geom_boxplot(aes(cage, value)) + 
+  facet_wrap(timepoint1 ~ ., ncol = 1) +
+  theme(axis.text.x = element_text(angle = 90))
 ```
-
-## Alpha diversity: tumor vs no tumor
-```{r}
-filt_alpha <- alpha_div %>% 
-  filter(timepoint %in% c('T2','T3')) %>%
-  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
-  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2)))
-
-ttest_res <- data.frame()
-for(measure in c('Observed', 'se.chao1', 'Simpson', 'Shannon')){
-  ttest_res %<>% bind_rows(data.frame(Measure=measure, pvalue=with(filt_alpha, t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
-}
-
-filt_alpha %>%
-  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon')) %>%
-  ggplot() +
-  geom_boxplot(aes(timepoint, value)) +
-  facet_grid(variable~.,scales = 'free_y')
-
-knitr::kable(ttest_res)
-```
-
-### Merge T1 and T2 and use them as baseline
-## Ordination: baseline vs T3
-```{r}
-ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T3'))) +
-  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
-  theme_bw()
-```
-
-## Ordination: baseline vs T4
-```{r}
-ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T4'))) +
-  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
-  theme_bw()
-```
-
-## Ordination: baseline vs T5
-```{r}
-ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T5'))) +
-  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
-  theme_bw()
-```
-
-## Ordination: baseline vs T6
-```{r}
-ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T6'))) +
-  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
-  theme_bw()
-```
-
-## Ordination: baseline vs T7
-```{r}
-ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T7'))) +
-  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
-  theme_bw()
-```
-
-## Ordination: T3 vs T7: Treated with Bifidum?
-```{r}
-ggplot(subset(ordwu, merge_timepoint %in% c('T3','T7'))) +
-  geom_point(aes(Axis.1, Axis.2, color=cond_all), size=3) +
-  scale_colour_discrete(name="Timepoint_Vaccine_Bifido") +
-  theme_bw() 
-```
-
-## Ordination: T3 vs T4
-```{r}
-metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T4','T3'),]
-distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T4','T3')), "wunifrac")
-permanova <- vegan::adonis(formula = distwu_2cat ~ timepoint + real_vaccine + timepoint:real_vaccine, data = metadata_2cat)
-
-subset(ordwu, timepoint %in% c('T4','T3')) %>%
-ggplot(aes(Axis.1, Axis.2)) +
-  geom_point(aes(color=timepoint, shape=real_vaccine), size=3, alpha=.7) +
-  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
-  scale_shape_manual(name = 'Pentatome?', values=c(15,19))+
-  theme_minimal() +
-  annotation_custom(grob = grid::textGrob(paste('Timepoint PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = .2, xmax = .2, ymin = .1, ymax = .1) +
-  annotation_custom(grob = grid::textGrob(paste('Pentatome PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[2], '\nR2=', signif(permanova$aov.tab$R2[2], 2)), 
-gp=grid::gpar(col="black", fontsize=10)),  xmin = -.1, xmax = -.1, ymin = -.08, ymax = -.08)
-```
-
-## Ordination: T3 vs T5
-```{r}
-metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T5','T3'),]
-distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T5','T3')), "wunifrac")
-permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
-
-subset(ordwu, timepoint %in% c('T5','T3')) %>%
-ggplot(aes(Axis.1, Axis.2)) +
-  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
-  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
-  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
-  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
-  theme_minimal() +
-  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .12, ymax = .12)
-```
-
-## Ordination: T3 vs T6
-```{r}
-metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T6','T3'),]
-distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T6','T3')), "wunifrac")
-permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
-
-subset(ordwu, timepoint %in% c('T6','T3')) %>%
-ggplot(aes(Axis.1, Axis.2)) +
-  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
-  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
-  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
-  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
-  theme_minimal() +
-  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
-```
-
-## Ordination: T3 vs T7
-```{r}
-metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
-distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
-permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
-
-subset(ordwu, timepoint %in% c('T7','T3')) %>%
-ggplot(aes(Axis.1, Axis.2)) +
-  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
-  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
-  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
-  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
-  theme_minimal() +
-  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
-```
-
-## Beta diversity between timepoints, T1 is the baseline
-```{r Beta diversity between timepoints t1}
-distances[distances$timepoint1 == 1 | distances$timepoint1 == 2, 'timepoint1'] <- 2
-distances[distances$timepoint2 == 1 | distances$timepoint2 == 2, 'timepoint2'] <- 2
-
-t1 <- distances %>% 
-  filter(topo1==topo2) %>%
-  filter(timepoint1==2) %>%
-  filter(timepoint1!=timepoint2) %>%
-  arrange(topo1)
-
-t.res <- data.frame()
-for(cond in unique(t1$condizione1)){
-  for(x in combn(c(3:7), 2, simplify = FALSE)){
-    pval.t = t.test(t1[t1$timepoint2==x[1] & t1$condizione1==cond,"value"], t1[t1$timepoint2==x[2] & t1$condizione1==cond ,"value"])$p.value
-    pval.w = wilcox.test(t1[t1$timepoint2==x[1] & t1$condizione1==cond,"value"], t1[t1$timepoint2==x[2] & t1$condizione1==cond ,"value"], paired = F)$p.value
-    if(pval.t < 0.05 | pval.w < 0.05)
-      t.res %<>% bind_rows(c(condizione1=cond, timepoint=as.numeric(x), p.value = str_c('t-test:',signif(pval.t, digits = 2),'U-test:', signif(pval.w, digits = 2), sep = ' ')))
-    }
-  }
-
-```
-
-```{r}
-knitr::kable(t.res)
-```
-
-```{r}
-
-t1 %>%
-  ggplot(aes(timepoint2, value, group = timepoint2)) +
-  stat_boxplot() +
-  scale_x_continuous(breaks = c(2:7)) +
-  # geom_signif(data=t.res,
-              # aes(xmin= as.numeric(timepoint1), xmax = as.numeric(timepoint2), annotations = p.value, y_position = 0.33),
-              # manual = TRUE) +
-  facet_grid(condizione1 ~ .)
-
-```
-
-## Beta diversity between timepoints, T(-1) is the baseline
-```{r Beta diversity between timepoints tn-1}
-tn1 <- distances %>% 
-  filter(topo1==topo2) %>%
-  filter(timepoint1==timepoint2-1) %>%
-  filter(timepoint1!=timepoint2) %>%
-  arrange(topo1)
-
-t.res <- data.frame()
-for(cond in unique(tn1$condizione1)){
-  for(x in combn(c(3:7), 2, simplify = FALSE)){
-      pval.t = t.test(tn1[tn1$timepoint2==x[1] & tn1$condizione1==cond,"value"], tn1[tn1$timepoint2==x[2] & tn1$condizione1==cond ,"value"])$p.value
-      pval.w = wilcox.test(tn1[tn1$timepoint2==x[1] & tn1$condizione1==cond,"value"], tn1[tn1$timepoint2==x[2] & tn1$condizione1==cond ,"value"], paired = F)$p.value
-      if(pval.t < 0.05 | pval.w < 0.05)
-        t.res %<>% bind_rows(c(condizione1=cond, timepoint=as.numeric(x), p.value = str_c('t-test:',signif(pval.t, digits = 2),'U-test:', signif(pval.w, digits = 2), sep = ' ')))
-  }
-}
-```
-
-```{r}
-knitr::kable(t.res)
-```
-
-```{r}
-tn1 %>%
-  ggplot(aes(timepoint2, value, group = timepoint2)) +
-  geom_boxplot() +
-  scale_x_continuous(breaks = c(2:7)) +
-  # geom_signif(data=t.res,
-              # aes(xmin= as.numeric(timepoint1), xmax = as.numeric(timepoint2), annotations = p.value, y_position = 0.33),
-              # manual = TRUE) +
-  facet_grid(condizione1 ~ .)
-```
-
-## Differential analysis
-```{r DA, cache=TRUE}
-threshold <- 0.05
-
-ggtopi_filt <- filter_taxa(gg_topi, function(x) mean(x) > 5e-1, TRUE)
-
-dds <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ cond_vac), parallel = TRUE, quiet = TRUE, fitType = 'parametric')
-
-dds_bif <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi, design = ~ cond_bif), parallel = TRUE, quiet = TRUE)
-dds1 <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ timepoint), parallel = TRUE, quiet = TRUE)
-
-dds_phylum <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi.phylum, design = ~ timepoint), parallel = TRUE, quiet = TRUE)
-
-```
-## T2 vs T3
-```{r DA T2 vs T3}
-res23 <- DESeq2::results(dds1, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
-res23.filtered <- res23[!is.na(res23$padj),]
-res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]
-res23.filtered <- res23.filtered[abs(res23.filtered$log2FoldChange) > 2,]
-
-res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
-  ggplot(aes(log2FoldChange, Genus)) + 
-  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-  geom_point(size=3, aes(color = log(baseMean))) + 
-  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-  theme_bw() +
-  ggtitle('T1 vs T2') +
-  scale_y_discrete(position = "right") +
-  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-        strip.text = element_text(size = 9),
-        axis.title = element_text(size = 9),
-        axis.title.y = element_blank(),
-        strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text.y = element_text(angle = 180)
-  )
-```
-
-## T2 vs T3 phylum level
-```{r DA T2 vs T3 phylum}
-res_phylum <- DESeq2::results(dds_phylum, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
-res_phylum.filtered <- res_phylum[!is.na(res_phylum$padj),]
-res_phylum.filtered <- res_phylum.filtered[res_phylum.filtered$padj < threshold,]
-# res_phylum.filtered <- res_phylum.filtered[abs(res_phylum.filtered$log2FoldChange) > 2,]
-
-res_phylum.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res_phylum.filtered),])) %>%
-  ggplot(aes(log2FoldChange, Phylum)) + 
-  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-  geom_point(size=3, aes(color = baseMean)) + 
-  theme_bw() +
-  scale_y_discrete(position = "right") +
-  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-        strip.text = element_text(size = 9),
-        axis.title = element_text(size = 9),
-        axis.title.y = element_blank(),
-        strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text.y = element_text(angle = 180)
-  )
-```
-
-## Check T1 vs T2
-```{r DA T1 vs T2}
-res12 <- DESeq2::results(dds1, contrast = c("timepoint","T1","T2"))
-res12.filtered <- res12[!is.na(res12$padj),]
-res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-res12.filtered %>% 
-  as.data.frame %>% 
-  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-  ggplot(aes(log2FoldChange, Genus)) + 
-  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-  geom_point(size=3, aes(color = baseMean)) + 
-  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-  theme_bw() +
-  ggtitle('Timepoint 1 vs Timepoint 2') +
-  scale_y_discrete(position = "right") +
-  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-        strip.text = element_text(size = 9),
-        axis.title = element_text(size = 9),
-        axis.title.y = element_blank(),
-        strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text.y = element_text(angle = 180)
-  )
-```
-
-```{r DA T3 vs T4}
-res23 <- DESeq2::results(dds1, contrast = c("timepoint","T3","T4"))
-res23.filtered <- res23[!is.na(res23$padj),]
-res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]
-
-res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
-  ggplot(aes(log2FoldChange, Genus)) + 
-  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-  geom_point(size=3, aes(color = baseMean)) + 
-  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-  theme_bw() +
-  ggtitle('Timepoint 3 vs Timepoint 4') +
-  scale_y_discrete(position = "right") +
-  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-        strip.text = element_text(size = 9),
-        axis.title = element_text(size = 9),
-        axis.title.y = element_blank(),
-        strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text.y = element_text(angle = 180)
-)
-```
-
-## Differential analysis Pentatome T7 vs not vaccinated T7
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"), independentFiltering = TRUE)
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = log10(baseMean))) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle('Pentatome T3 vs T7') +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-## Differential analysis Pentatome T3 VS T7
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T3_NO"))
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = log10(baseMean))) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle('Pentatome T3 vs T7') +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-```{r}
-res12.filtered %>% 
-  as.data.frame %>% 
-  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-  filter(is.na(Phylum)) %>%  knitr::kable()
-```
-
-## Differential analysis Empty vescicles T3 VS T7
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_NO","T3_NO"))
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = baseMean)) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle('Empty T3vsT7') +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-## Differential analysis Pentatoma vs empty vescicles T7
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = baseMean)) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-## Differential analysis Pentatoma vs empty vescicles T3
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T3_YES","T3_NO"))
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-  timepoint_number <- stringr::str_sub(timepoint,17,17)
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = baseMean)) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-## Differential analysis Pentatome vs No-pentatome at different timepoints
-```{r}
-for (timepoint in c("merge_timepointT3.pentatomePBS","merge_timepointT4.pentatomePBS","merge_timepointT5.pentatomePBS","merge_timepointT6.pentatomePBS","merge_timepointT7.pentatomePBS")) {
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T6_Pentatoma","T6_Empty"), )
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-  timepoint_number <- stringr::str_sub(timepoint,17,17)
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = baseMean)) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-}
-```
-
-<!-- ## Differential analysis Bifido vs No-bifido at different timepoints -->
-<!-- ```{r} -->
-<!-- for (timepoint in c("bifidoPBS.merge_timepointT3","bifidoPBS.merge_timepointT4","bifidoPBS.merge_timepointT5","bifidoPBS.merge_timepointT6","bifidoPBS.merge_timepointT7")) { -->
-<!--   res12 <- DESeq2::results(dds2, contrast = list(c("bifido_PBS_vs_Bifido" , timepoint))) -->
-<!--   res12.filtered <- res12[!is.na(res12$padj),] -->
-<!--   res12.filtered <- res12.filtered[res12.filtered$padj < threshold,] -->
-<!--   timepoint_number <- stringr::str_sub(timepoint,-1) -->
-
-<!--   plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>% -->
-<!--     ggplot(aes(log2FoldChange, Genus)) + -->
-<!--     geom_vline(xintercept = 0.0, color = "gray", size = 0.5) + -->
-<!--     geom_point(size=3, aes(color = baseMean)) + -->
-<!--     facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+ -->
-<!--     theme_bw() + -->
-<!--     ggtitle(paste('Bifido vs no bifido at timepoint', timepoint_number, sep = ' ')) + -->
-<!--     scale_y_discrete(position = "right") + -->
-<!--     theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11), -->
-<!--           strip.text = element_text(size = 9), -->
-<!--           axis.title = element_text(size = 9), -->
-<!--           axis.title.y = element_blank(), -->
-<!--           strip.background = element_rect(colour = "transparent", fill = "transparent"), -->
-<!--           strip.text.y = element_text(angle = 180) -->
-<!--     )) -->
-<!-- } -->
-<!-- ``` -->
-
-# PERMANOVA
-```{r}
-metadata <- as(sample_data(gg_topi), "data.frame")
-permanova.res <- data.frame()
-distwu <- phyloseq::distance(gg_topi, "wunifrac")
-
-all_permanova <- as.data.frame(vegan::adonis(formula = distwu ~ merge_timepoint*bifido*pentatome, data = metadata)$aov.tab)
-```
-
-```{r}
-knitr::kable(all_permanova)
-```
-
-```{r}
-for(s in combn(levels(metadata$merge_timepoint),2, simplify = FALSE)){
-  metadata_2cat <- metadata[metadata$merge_timepoint %in% s,]
-  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% s), "wunifrac")
-  permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
-  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
-}
-```
-
-```{r}
-for(s in combn(levels(metadata$pentatome),2, simplify = FALSE)){
-  metadata_2cat <- metadata[metadata$pentatome %in% s,]
-  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, pentatome %in% s), "wunifrac")
-  permanova <- vegan::adonis(formula = distwu_2cat ~ pentatome, data = metadata_2cat)
-  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
-}
-```
-
-```{r}
-for(s in combn(levels(metadata$bifido),2, simplify = FALSE)){
-  metadata_2cat <- metadata[metadata$bifido %in% s,]
-  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, bifido %in% s), "wunifrac")
-  permanova <- vegan::adonis(formula = distwu_2cat ~ bifido, data = metadata_2cat)
-  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
-}
-```
-
-```{r}
-knitr::kable(permanova.res)
-```
-
-## PERMANOVA all vs all
-```{r}
-permanova.res <- data.frame()
-for (s in combn(levels(sample_data(gg_topi)$cond_all),2, simplify = FALSE)) {
-  metadata_2cat <- data.frame(sample_data(gg_topi))[sample_data(gg_topi)$cond_all %in% s,]
-  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, cond_all %in% s), "wunifrac")
-  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat)
-  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
-}
-```
-
-```{r}
-knitr::kable(permanova.res)
-```
-
-## (T5+T6+T7) vs baseline
-```{r}
-metadata_tub <- metadata[metadata$merge_timepoint %in% c('baseline','T5','T6','T7'),]
-metadata_tub$merge_timepoint <- as.character(metadata_tub$merge_timepoint)
-metadata_tub[which(metadata_tub$merge_timepoint!='baseline'), 'merge_timepoint'] <- 'Timm'
-distwu_tub <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% c('baseline','T5','T6','T7')), "wunifrac")
-permanova <- vegan::adonis(formula = distwu_tub ~ merge_timepoint, data = metadata_tub)
-```
-
-```{r}
-knitr::kable(permanova$aov.tab)
-```
-
-## Effect of bifido in T4 vs T5
-```{r}
-metadata_t45 <- metadata[metadata$merge_timepoint %in% c('T4','T5'),]
-distwu_t45 <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% c('T4','T5')), "wunifrac")
-permanova <- vegan::adonis(formula = distwu_t45 ~ merge_timepoint*bifido, data = metadata_t45)
-```
-
-```{r}
-knitr::kable(permanova$aov.tab)
-```
-
-## Beta diversity T3 vs T4
-```{r}
-t1 <- distances %>% 
-  dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==3) %>%
-  dplyr::filter(timepoint2==4) %>%
-  dplyr::arrange(topo1)
-
-t1 %>%
-  ggplot(aes(pentatome1, value)) +
-  stat_boxplot() +
-  stat_signif()
-```
-
-## Beta diversity T3 vs T7
-```{r}
-t37 <- distances %>% 
-  dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==3) %>%
-  dplyr::filter(timepoint2==7) %>%
-  dplyr::arrange(topo1)
-
-t37 %>%
-  ggplot(aes(condizione1, value)) +
-  stat_boxplot()
-```
-
-## Correlation beta-diversity T2-TX, tumor mass
-```{r corr tumor wunifrac}
-distances %>% 
-  dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==2) %>%
-  dplyr::filter(timepoint2>4) %>%
-  dplyr::select(cols, timepoint2, value) %>%
-  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
-  ggplot(aes(value,tumor_mass)) +
-  geom_point() + 
-  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
-  geom_smooth(method = 'lm')
-```
-
-```{r}
-wunif_ds <- distances %>% 
-  dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==2) %>%
-  dplyr::filter(timepoint2>4) %>%
-  dplyr::select(cols, timepoint2, value) %>%
-  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))
-
-ds_corr_wunif <- data.frame()
-for(x in unique(wunif_ds$timepoint2)){
-  for(y in levels(wunif_ds$condizione)){
-    tmp_res <- with(wunif_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
-    ds_corr_wunif %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
-  }
-}
-
-knitr::kable(ds_corr_wunif)
-```
-
-```{r corr tumor braycurtis}
-distancesbc %>% 
-  dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==2) %>%
-  dplyr::filter(timepoint2>4) %>%
-  dplyr::select(cols, timepoint2, value) %>%
-  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
-  ggplot(aes(value,tumor_mass)) +
-  geom_point() + 
-  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
-  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
-  geom_smooth(method = 'lm')
-```
-
-```{r}
-brayc_ds <- distancesbc %>% 
-  dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==2) %>%
-  dplyr::filter(timepoint2>4) %>%
-  dplyr::select(cols, timepoint2, value) %>%
-  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))
-
-ds_corr_braycf <- data.frame()
-for(x in unique(brayc_ds$timepoint2)){
-  for(y in levels(brayc_ds$condizione)){
-    tmp_res <- with(brayc_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
-    ds_corr_braycf %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
-  }
-}
-
-knitr::kable(ds_corr_braycf)
-```
-
-```{r corr tumor jaccard}
-distancesjac %>% 
-  dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==2) %>%
-  dplyr::filter(timepoint2>4) %>%
-  dplyr::select(cols, timepoint2, value) %>%
-  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
-  ggplot(aes(value,tumor_mass)) +
-  geom_point() + 
-  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
-  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
-  geom_smooth(method = 'lm')
-```
-
-```{r}
-jac_ds <- distancesjac %>% 
-  dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==2) %>%
-  dplyr::filter(timepoint2>4) %>%
-  dplyr::select(cols, timepoint2, value) %>%
-  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))
-
-ds_corr_jac <- data.frame()
-for(x in unique(jac_ds$timepoint2)){
-  for(y in levels(jac_ds$condizione)){
-    tmp_res <- with(jac_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
-    ds_corr_jac %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
-  }
-}
-
-knitr::kable(ds_corr_jac)
-```
-
-# Tracking OTU
-```{r}
-counts <- data.frame(as.data.frame(otu_table(gg_topi)[c('EU622777.1.1487','EU504401.1.1405'),] %>% t)) %>% tibble::rownames_to_column('sampleid')
-colnames(counts) <- c('sampleid',"Lachnospiraceae_NK4A136","Muribaculaceae")
-counts$t <- sapply(as.character(counts$sampleid), function(x) strsplit(x,'-')[[1]][4])
-counts$condizione <- sapply(as.character(counts$sampleid), function(x) strsplit(x,'-')[[1]][2])
-
-reshape2::melt(counts, id.vars = c('t', 'condizione','sampleid')) %>%
-  ggplot() +
-  stat_boxplot(aes(t, value)) + 
-  facet_grid(variable ~ condizione, scales = "free_x", space="free_y")
-```

### 2018-11-19 da_test.Rmd
T7 no 24/7

---
title: "Differential Analysis"
output: 
  html_document: 
    number_sections: yes
    theme: cosmo
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(phyloseq)
library(magrittr)
library(tidyverse)
library(reshape2)
library(ggsignif)
library(DESeq2)

```

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/otu_table_mc10_w_tax_rdp.biom',
                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.tre',
                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.fna',
                        refseqFunction = parse_taxonomy_default
                      )

mappingfile <- read.delim('/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/mappingfile', row.names = 1, check.names = FALSE)
mappingfile$Description %<>% str_replace_all(.,'_','-')
mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
rownames(mappingfile) <- mappingfile$Description

sample_data(gg_topi) <- mappingfile
sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)

sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])

sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))

sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)

colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")

tax_table_filtered <- tax_table(gg_topi) %>%
  as.data.frame %>% 
  mutate_all( .funs = function(x) stringr:::str_replace_all(x,"D_.__",""))

rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
tax_table(gg_topi) <-  as.matrix(tax_table_filtered)

unass_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(Domain == 'Unassigned') %>%
                select(OTU)

unknown_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(is.na(Genus)) %>%
                select(OTU)

unknown_family <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(is.na(Family)) %>%
                select(OTU)

tax_table(gg_topi)[unass_otus$OTU, ] <- rep.int('Unassigned', 7)
tax_table(gg_topi)[unknown_otus$OTU, 'Genus'] <- 'Unknown'
tax_table(gg_topi)[unknown_family$OTU, 'Family'] <- 'Unknown'

gg_topi.relab <- transform_sample_counts(gg_topi, function(x) x/sum(x))

gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

gg_topi.family <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.family.relab <- transform_sample_counts(gg_topi.family, function(x) x/sum(x))
```

# Differential Analysis with DESeq2
```{r, cache=TRUE}
threshold <- 0.05

ggtopi_filt <- filter_taxa(gg_topi, function(x) mean(x) > 5e-1, TRUE)

pds <- phyloseq_to_deseq2(ggtopi_filt, design = ~timepoint)
dds <- DESeq2::DESeq(pds, parallel = TRUE, quiet = TRUE)

pds_genus <- phyloseq_to_deseq2(gg_topi.genus, design = ~timepoint)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE)

pds_family <- phyloseq_to_deseq2(gg_topi.family, design = ~timepoint)
dds_family <- DESeq2::DESeq(pds_family, parallel = TRUE, quiet = TRUE)
```

## T2 vs T3: Tumor vs no tumor injected
```{r}
res <- DESeq2::results(dds, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
res.filt <- res[!is.na(res$padj),]
res.filt <- res.filt[res.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_filt)[rownames(res.filt),])) %>%
  filter(baseMean >= 10) %>%
  ggplot(aes(log2FoldChange, interaction(Family,Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot abundances
```{r}
  t23_samples_idx <- grep('T2|T3', sample_names(gg_topi.relab), perl = TRUE)

  otu_ab_t <- otu_table(gg_topi.relab)[rownames(res.filt), t23_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t %>% 
    filter(OTU %in% (res.filt[res.filt$baseMean >= 10,] %>% rownames)) %>%
    bind_cols(Genus = as.matrix(tax_table(gg_topi.relab))[.$OTU,'Genus'],
              Family = as.matrix(tax_table(gg_topi.relab))[.$OTU,'Family']) %>%
    ggplot() +
      geom_boxplot(aes(interaction(Family, Genus, OTU), Abundance, fill=timepoint)) + 
      coord_flip()
```

## T2 vs T3: Tumor vs no tumor injected at Genus level
```{r}
res_genus <- DESeq2::results(dds_genus, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
res.filt_genus <- res_genus[!is.na(res_genus$padj),]
res.filt_genus <- res.filt_genus[res.filt_genus$padj<0.05,]
```

### Dotplot log2FC vs Genus
```{r}
res.filt_genus %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi.genus)[rownames(res.filt_genus),-7])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) +
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of Genus abundances of differenitally abundant Genera
```{r}
  t23_samples_idx_genus <- grep('T2|T3', sample_names(gg_topi.genus.relab), perl = TRUE)

  genus_ab_t <- otu_table(gg_topi.genus.relab)[rownames(res.filt_genus), t23_samples_idx_genus] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  genus_ab_t %>% bind_cols(data.frame(tax_table(gg_topi.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## T2 vs T3: Tumor vs no tumor injected at Family level
```{r}
res_family <- DESeq2::results(dds_family, contrast = c('timepoint', 'T2','T3'), alpha = 0.05) 
res.filt_family <- res_family[!is.na(res_family$padj),]
res.filt_family <- res.filt_family[res.filt_family$padj<0.05,]
```

### Dotplot log2FC vs Family
```{r}
res.filt_family %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi.family)[rownames(res.filt_family),-7])) %>%
  ggplot(aes(log2FoldChange, Family)) +
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot abundances
```{r}
  t23_samples_idx_family <- grep('T2|T3', sample_names(gg_topi.family.relab), perl = TRUE)

  family_ab_t <- otu_table(gg_topi.family.relab)[rownames(res.filt_family), t23_samples_idx_family] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  family_ab_t %>% bind_cols(data.frame(tax_table(gg_topi.family.relab)[family_ab_t$OTU,c('Phylum','Family')])) %>%
  ggplot() +
    geom_boxplot(aes(Family, Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## T2 vs T1: Tumor vs no tumor injected
```{r}
res21 <- DESeq2::results(dds, contrast = c('timepoint', 'T2','T1'), alpha = 0.05)
res21.filt <- res21[!is.na(res21$padj),]
res21.filt <- res21.filt[res21.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res21.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_filt)[rownames(res21.filt),])) %>%
  filter(baseMean >= 10) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot abundances
```{r}
  t12_samples_idx <- grep('T2|T3', sample_names(ggtopi_filt), perl = TRUE)

  otu_ab_t12 <- otu_table(gg_topi.relab)[rownames(res21.filt), t12_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t12 %>% bind_cols(data.frame(tax_table(ggtopi_filt)[otu_ab_t12$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

# T3 vs T7
```{r, cache=TRUE}
ggtopi_subs <-subset_samples(gg_topi, timepoint %in% c('T3','T7'))
ggtopi_subs.genus <- tax_glom(ggtopi_subs,taxrank = "Genus")
ggtopi_subs.genus.relab <- transform_sample_counts(ggtopi_subs.genus, function(x) x/sum(x))
mmtx <- model.matrix(object = ~timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, data = data.frame(ggtopi_subs@sam_data))
idx <- which(apply(mmtx, 2, function(x) all(x==0)))
mmtx <- mmtx[,-idx]

pds_37 <- phyloseq_to_deseq2(ggtopi_subs, design = ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_37 <- DESeq2::DESeq(pds_37, parallel = TRUE, quiet = TRUE, full = mmtx)

pds_genus <- phyloseq_to_deseq2(ggtopi_subs.genus, design =  ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE, full = mmtx)
```

## What is altered at T7 vs T3?
```{r}
res37 <- DESeq2::results(dds_37, name = "timepointT7", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs)[rownames(res37.filt),])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of OTUs abundances of differenitally abundant OTUs
```{r}
  t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_subs), perl = TRUE)

  otu_ab_t37 <- otu_table(ggtopi_subs)[rownames(res37.filt), t37_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t37 %>% bind_cols(data.frame(tax_table(ggtopi_filt)[otu_ab_t37$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## What is altered at T7 vs T3 at genus levels?
```{r}
res37 <- DESeq2::results(dds_genus, name = "timepointT7", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs Genus
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs.genus)[rownames(res37.filt),])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of Genus abundances of differenitally abundant Genera
```{r}
  genus_ab_t <- otu_table(ggtopi_subs.genus.relab)[rownames(res37.filt), ] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  genus_ab_t %>% bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## What is altered at T7 after vaccination and Bifidobacterium gavage?
```{r}
res37 <- DESeq2::results(dds_37, name = "timepointT7.real_vaccineYES.real_bifidoYES", alpha = 0.05) 
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs)[rownames(res37.filt),])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of OTUs abundances of differenitally abundant OTUs
```{r}
  t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_filt), perl = TRUE)
  tax_table(gg_topi.relab)[rownames(res37.filt),]

  otu_ab_t37 <- otu_table(gg_topi.relab)[rownames(res37.filt), t37_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t37 %>% bind_cols(data.frame(tax_table(ggtopi_filt)[otu_ab_t37$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## What is altered at T7 after vaccination only?
```{r}
res37 <- DESeq2::results(dds_37, name = "timepointT7.real_vaccineYES", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs)[rownames(res37.filt),])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of OTUs abundances of differenitally abundant OTUs
```{r}
  t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_filt), perl = TRUE)

  otu_ab_t37 <- otu_table(gg_topi.relab)[rownames(res37.filt), t37_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t37 %>% bind_cols(data.frame(tax_table(ggtopi_filt)[otu_ab_t37$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## What is altered at T7 after bifido gavage?
```{r}
res37 <- DESeq2::results(dds_37, name = "timepointT7.real_bifidoYES", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

NOTHING

# Removal of all Bacteroidales 24/7
```{r}
gg_topi_no247 <- subset_taxa(gg_topi, !Family %in% c("Bacteroidales S24-7 group"))
```

# T3 vs T7 without 24/7
```{r, cache=TRUE}
ggtopi_subs <-subset_samples(gg_topi_no247, timepoint %in% c('T3','T7'))
ggtopi_subs.genus <- tax_glom(ggtopi_subs,taxrank = "Genus")
ggtopi_subs.genus.relab <- transform_sample_counts(ggtopi_subs.genus, function(x) x/sum(x))
mmtx <- model.matrix(object = ~timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, data = data.frame(ggtopi_subs@sam_data))
idx <- which(apply(mmtx, 2, function(x) all(x==0)))
mmtx <- mmtx[,-idx]

pds_37 <- phyloseq_to_deseq2(ggtopi_subs, design = ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_37 <- DESeq2::DESeq(pds_37, parallel = TRUE, quiet = TRUE, full = mmtx)

pds_genus <- phyloseq_to_deseq2(ggtopi_subs.genus, design =  ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE, full = mmtx)
```

## What is altered at T7 vs T3 at genus levels?
```{r}
res37 <- DESeq2::results(dds_genus, name = "timepointT7", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs Genus
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs.genus)[rownames(res37.filt),-7])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### 2018-11-19 gg_topi.Rmd
T7 no 24/7

@@
 ---
-title: "GG_topi2"
+title: "GG_topi"
 author: "Francesco Beghini"
@@
   html_document: 
+    df_print: kable
+    self_contained: no
     theme: cerulean
 editor_options: 
-  chunk_output_type: console
+  chunk_output_type: inline
 ---
@@
 ```{r load qiime, include=FALSE, cache=TRUE}
-gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/otu_table_mc10_w_tax.biom',
-                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/rep_set.tre',
-                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus_silva/rep_set.fna',
+gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/otu_table_mc10_w_tax_rdp.biom',
+                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.tre',
+                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.fna',
                         refseqFunction = parse_taxonomy_default
@@
 tax_table(gg_topi) <-  as.matrix(tax_table_filtered)
+
+unass_otus <- tax_table(gg_topi) %>%
+                data.frame %>%
+                tibble::rownames_to_column('OTU') %>%
+                filter(Domain == 'Unassigned') %>%
+                select(OTU)
+
+unknown_otus <- tax_table(gg_topi) %>%
+                data.frame %>%
+                tibble::rownames_to_column('OTU') %>%
+                filter(is.na(Genus)) %>%
+                select(OTU)
+
+tax_table(gg_topi)[unass_otus$OTU, ] <- rep.int('Unassigned', 7)
+tax_table(gg_topi)[unknown_otus$OTU, 'Genus'] <- 'Unknown'
 
@@
 # Beta diversity
-```{r Beta diversity, cache=TRUE}
+```{r Beta diversity, cache=FALSE}
 distwu <- phyloseq::distance(gg_topi, "wunifrac")
@@
 ```{r Beta diversity between timepoints t1}
-distances[distances$timepoint1 == 1 | distances$timepoint1 == 2, 'timepoint1'] <- 2
-distances[distances$timepoint2 == 1 | distances$timepoint2 == 2, 'timepoint2'] <- 2
+distances_tmp <- distances
+distances_tmp[distances_tmp$timepoint1 == 1 | distances_tmp$timepoint1 == 2, 'timepoint1'] <- 2
+distances_tmp[distances_tmp$timepoint2 == 1 | distances_tmp$timepoint2 == 2, 'timepoint2'] <- 2
 
-t1 <- distances %>% 
+t1 <- distances_tmp %>% 
   filter(topo1==topo2) %>%
@@
 
-## Differential analysis
-```{r DA, cache=TRUE}
-threshold <- 0.05
-
-ggtopi_filt <- filter_taxa(gg_topi, function(x) mean(x) > 5e-1, TRUE)
-
-dds <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ cond_vac), parallel = TRUE, quiet = TRUE, fitType = 'parametric')
-
-mmtx <- model.matrix(object = ~timepoint*real_vaccine*real_bifido, data = data.frame(gg_topi@sam_data))
-idx <- which(apply(mmtx, 2, function(x) all(x==0)))
-mmtx <- mmtx[,-idx]
-
-rmmtx <- model.matrix(object = ~ timepoint, data = data.frame(gg_topi@sam_data))
-
-pds <- phyloseq_to_deseq2(ggtopi_filt, design = ~timepoint*real_vaccine*real_bifido, ignoreRank = TRUE)
-
-dds <- DESeq2::DESeq(pds, full = mmtx, reduced = rmmtx, test = 'LRT')
-
-dds_bif <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi, design = ~ cond_bif), parallel = TRUE, quiet = TRUE)
-dds1 <- DESeq2::DESeq(phyloseq_to_deseq2(ggtopi_filt, design = ~ timepoint), parallel = TRUE, quiet = TRUE)
-
-dds_phylum <- DESeq2::DESeq(phyloseq_to_deseq2(gg_topi.phylum, design = ~ timepoint), parallel = TRUE, quiet = TRUE)
-
-```
-## T2 vs T3
-```{r DA T2 vs T3}
-res23 <- DESeq2::results(dds1, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
-res23.filtered <- res23[!is.na(res23$padj),]
-res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]
-res23.filtered <- res23.filtered[abs(res23.filtered$log2FoldChange) > 2,]
-
-res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
-  ggplot(aes(log2FoldChange, Genus)) + 
-  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-  geom_point(size=3, aes(color = log(baseMean))) + 
-  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-  theme_bw() +
-  ggtitle('T1 vs T2') +
-  scale_y_discrete(position = "right") +
-  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-        strip.text = element_text(size = 9),
-        axis.title = element_text(size = 9),
-        axis.title.y = element_blank(),
-        strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text.y = element_text(angle = 180)
-  )
-```
-
-## T2 vs T3 phylum level
-```{r DA T2 vs T3 phylum}
-res_phylum <- DESeq2::results(dds_phylum, contrast = c("timepoint","T2","T3"), alpha = 0.05, format = "DataFrame")
-res_phylum.filtered <- res_phylum[!is.na(res_phylum$padj),]
-res_phylum.filtered <- res_phylum.filtered[res_phylum.filtered$padj < threshold,]
-# res_phylum.filtered <- res_phylum.filtered[abs(res_phylum.filtered$log2FoldChange) > 2,]
-
-res_phylum.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res_phylum.filtered),])) %>%
-  ggplot(aes(log2FoldChange, Phylum)) + 
-  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-  geom_point(size=3, aes(color = baseMean)) + 
-  theme_bw() +
-  scale_y_discrete(position = "right") +
-  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-        strip.text = element_text(size = 9),
-        axis.title = element_text(size = 9),
-        axis.title.y = element_blank(),
-        strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text.y = element_text(angle = 180)
-  )
-```
-
-## Check T1 vs T2
-```{r DA T1 vs T2}
-res12 <- DESeq2::results(dds1, contrast = c("timepoint","T1","T2"))
-res12.filtered <- res12[!is.na(res12$padj),]
-res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-res12.filtered %>% 
-  as.data.frame %>% 
-  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-  ggplot(aes(log2FoldChange, Genus)) + 
-  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-  geom_point(size=3, aes(color = baseMean)) + 
-  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-  theme_bw() +
-  ggtitle('Timepoint 1 vs Timepoint 2') +
-  scale_y_discrete(position = "right") +
-  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-        strip.text = element_text(size = 9),
-        axis.title = element_text(size = 9),
-        axis.title.y = element_blank(),
-        strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text.y = element_text(angle = 180)
-  )
-```
-
-```{r DA T3 vs T4}
-res23 <- DESeq2::results(dds1, contrast = c("timepoint","T3","T4"))
-res23.filtered <- res23[!is.na(res23$padj),]
-res23.filtered <- res23.filtered[res23.filtered$padj < threshold,]
-
-res23.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res23.filtered),])) %>%
-  ggplot(aes(log2FoldChange, Genus)) + 
-  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-  geom_point(size=3, aes(color = baseMean)) + 
-  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-  theme_bw() +
-  ggtitle('Timepoint 3 vs Timepoint 4') +
-  scale_y_discrete(position = "right") +
-  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-        strip.text = element_text(size = 9),
-        axis.title = element_text(size = 9),
-        axis.title.y = element_blank(),
-        strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text.y = element_text(angle = 180)
-)
-```
-
-## Differential analysis Pentatome T7 vs not vaccinated T7
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
-  res12 <- DESeq2::results(dds)
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = log10(baseMean))) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle('Pentatome T3 vs T7') +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-## Differential analysis Pentatome T3 VS T7
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T3_NO"))
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = log10(baseMean))) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle('Pentatome T3 vs T7') +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-```{r}
-res12.filtered %>% 
-  as.data.frame %>% 
-  cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-  filter(is.na(Phylum)) %>%  knitr::kable()
-```
-
-## Differential analysis Empty vescicles T3 VS T7
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_NO","T3_NO"))
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = baseMean)) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle('Empty T3vsT7') +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-## Differential analysis Pentatoma vs empty vescicles T7
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T7_YES","T7_NO"))
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = baseMean)) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-## Differential analysis Pentatoma vs empty vescicles T3
-```{r}
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T3_YES","T3_NO"))
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-  timepoint_number <- stringr::str_sub(timepoint,17,17)
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = baseMean)) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-
-```
-
-## Differential analysis Pentatome vs No-pentatome at different timepoints
-```{r}
-for (timepoint in c("merge_timepointT3.pentatomePBS","merge_timepointT4.pentatomePBS","merge_timepointT5.pentatomePBS","merge_timepointT6.pentatomePBS","merge_timepointT7.pentatomePBS")) {
-  res12 <- DESeq2::results(dds, contrast = c("cond_vac","T6_Pentatoma","T6_Empty"), )
-  res12.filtered <- res12[!is.na(res12$padj),]
-  res12.filtered <- res12.filtered[res12.filtered$padj < threshold,]
-  timepoint_number <- stringr::str_sub(timepoint,17,17)
-
-  plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>%
-    ggplot(aes(log2FoldChange, Genus)) +
-    geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
-    geom_point(size=3, aes(color = baseMean)) +
-    facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
-    theme_bw() +
-    ggtitle(paste('PBS vs Pentatoma at timepoint', timepoint_number, sep = ' ')) +
-    scale_y_discrete(position = "right") +
-    theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
-          strip.text = element_text(size = 9),
-          axis.title = element_text(size = 9),
-          axis.title.y = element_blank(),
-          strip.background = element_rect(colour = "transparent", fill = "transparent"),
-          strip.text.y = element_text(angle = 180)
-    ))
-}
-```
-
-<!-- ## Differential analysis Bifido vs No-bifido at different timepoints -->
-<!-- ```{r} -->
-<!-- for (timepoint in c("bifidoPBS.merge_timepointT3","bifidoPBS.merge_timepointT4","bifidoPBS.merge_timepointT5","bifidoPBS.merge_timepointT6","bifidoPBS.merge_timepointT7")) { -->
-<!--   res12 <- DESeq2::results(dds2, contrast = list(c("bifido_PBS_vs_Bifido" , timepoint))) -->
-<!--   res12.filtered <- res12[!is.na(res12$padj),] -->
-<!--   res12.filtered <- res12.filtered[res12.filtered$padj < threshold,] -->
-<!--   timepoint_number <- stringr::str_sub(timepoint,-1) -->
-
-<!--   plot(res12.filtered %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi)[rownames(res12.filtered),])) %>% -->
-<!--     ggplot(aes(log2FoldChange, Genus)) + -->
-<!--     geom_vline(xintercept = 0.0, color = "gray", size = 0.5) + -->
-<!--     geom_point(size=3, aes(color = baseMean)) + -->
-<!--     facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+ -->
-<!--     theme_bw() + -->
-<!--     ggtitle(paste('Bifido vs no bifido at timepoint', timepoint_number, sep = ' ')) + -->
-<!--     scale_y_discrete(position = "right") + -->
-<!--     theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11), -->
-<!--           strip.text = element_text(size = 9), -->
-<!--           axis.title = element_text(size = 9), -->
-<!--           axis.title.y = element_blank(), -->
-<!--           strip.background = element_rect(colour = "transparent", fill = "transparent"), -->
-<!--           strip.text.y = element_text(angle = 180) -->
-<!--     )) -->
-<!-- } -->
-<!-- ``` -->
-
 # PERMANOVA
-```{r}
+```{r, cache=TRUE}
 metadata <- as(sample_data(gg_topi), "data.frame")
@@
 
-```{r}
+```{r, cache=TRUE}
 for(s in combn(levels(metadata$merge_timepoint),2, simplify = FALSE)){
@@
 
-```{r}
+```{r, cache=TRUE}
 for(s in combn(levels(metadata$pentatome),2, simplify = FALSE)){
@@
 
-```{r}
+```{r, cache=TRUE}
 for(s in combn(levels(metadata$bifido),2, simplify = FALSE)){
@@
 ## PERMANOVA all vs all
-```{r}
+```{r, cache=TRUE}
 permanova.res <- data.frame()
@@
   distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, cond_all %in% s), "wunifrac")
-  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat)
+  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat, parallel = 2)
   permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
@@
 
-## Correlation beta-diversity T2-TX, tumor mass
+## Correlation beta-diversity T3-TX, tumor mass
 ```{r corr tumor wunifrac}
@@
   dplyr::filter(topo1==topo2) %>%
-  dplyr::filter(timepoint1==2) %>%
+  dplyr::filter(timepoint1==3) %>%
   dplyr::filter(timepoint2>4) %>%
@@
   geom_point() +
-  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
   facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
@@
   geom_point() + 
-  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
   facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
@@
 
-# Tracking OTU
+```{r, include=FALSE, eval=FALSE}
+otus <- data.frame(otu_table(gg_topi.genus.relab), check.names = FALSE) %>% 
+  tibble::rownames_to_column('OTU') %>% 
+  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+  tibble::add_column(., timepoint = stringr::str_sub(.$SampleID, 12, 13)) %>%
+  tibble::add_column(., mouseID = stringr::str_sub(.$SampleID, 8, 10))
+
+wb<-createWorkbook(type="xlsx")
+for(mID in sort(unique(otus$mouseID))){
+  tmp <- otus %>%
+    filter(mouseID==mID) %>%
+    reshape2::dcast(OTU ~ timepoint, value.var = 'Abundance') %>%
+    left_join(data.frame(tax_table(gg_topi.genus.relab)[,-7], check.names = FALSE) %>% tibble::rownames_to_column('OTU'))
+  sheet <- createSheet(wb, sheetName = mID)
+  addDataFrame(tmp, sheet)
+}
+saveWorkbook(wb, "otutable_topi.xlsx")
+```
+
+# Distribution of distances between conditions per timepoint
 ```{r}
-counts <- data.frame(as.data.frame(otu_table(gg_topi)[c('EU622777.1.1487','EU504401.1.1405'),] %>% t)) %>% tibble::rownames_to_column('sampleid')
-colnames(counts) <- c('sampleid',"Lachnospiraceae_NK4A136","Muribaculaceae")
-counts$t <- sapply(as.character(counts$sampleid), function(x) strsplit(x,'-')[[1]][4])
-counts$condizione <- sapply(as.character(counts$sampleid), function(x) strsplit(x,'-')[[1]][2])
+dct <- distances %>%
+  filter(timepoint1 == timepoint2) %>%
+  select(timepoint1, condizione1, condizione2, value) %>%
+  dplyr::rename(timepoint = timepoint1) %>% 
+  group_by(timepoint, condizione1, condizione2) %>%
+  summarise(mediandistance=median(value)) 
+```
+```{r include=TRUE,  results = "asis"}
+for(t in unique(dct$timepoint)){
+  print(dct %>%
+    filter(timepoint == t) %>% 
+    dplyr::rename(condizione = condizione2) %>%
+    spread(condizione1, mediandistance) %>%
+    knitr::kable())
 
-reshape2::melt(counts, id.vars = c('t', 'condizione','sampleid')) %>%
-  ggplot() +
-  stat_boxplot(aes(t, value)) + 
-  facet_grid(variable ~ condizione, scales = "free_x", space="free_y")
+}
 ```
+
+# Distribution of distances between timepoints per conditions
+```{r}
+dtc <- distances %>%
+  filter(condizione1 == condizione2) %>%
+  dplyr::rename(condizione = condizione1) %>% 
+  group_by(condizione, timepoint1, timepoint2) %>%
+  summarise(mediandistance=median(value)) 
+```
+
+```{r include=TRUE,  results = "asis"}
+for(c in unique(dtc$condizione)){
+  print(dtc %>%
+    filter(condizione == c) %>% 
+    dplyr::rename(timepoint = timepoint1) %>%
+    spread(timepoint2, mediandistance) %>%
+    knitr::kable())
+    cat("\n")
+}
+```

### 2018-11-19 da_test.Rmd
Stratify per condition

@@
 ggtopi_subs <-subset_samples(gg_topi, timepoint %in% c('T3','T7'))
+ggtopi_subs.relab <- transform_sample_counts(ggtopi_subs, function(x) x/sum(x))
 ggtopi_subs.genus <- tax_glom(ggtopi_subs,taxrank = "Genus")
@@
 
-  otu_ab_t37 <- otu_table(ggtopi_subs)[rownames(res37.filt), t37_samples_idx] %>%
+  otu_ab_t37 <- otu_table(ggtopi_subs.relab)[rownames(res37.filt), t37_samples_idx] %>%
               data.frame %>%
@@
 
-  otu_ab_t37 %>% bind_cols(data.frame(tax_table(ggtopi_filt)[otu_ab_t37$OTU,c('Phylum','Family','Genus')])) %>%
+  otu_ab_t37 %>% bind_cols(data.frame(tax_table(ggtopi_subs.relab)[otu_ab_t37$OTU,c('Phylum','Family','Genus')])) %>%
   ggplot() +
@@
 res37.filt %>% as.data.frame %>% 
-  cbind(as.matrix(tax_table(ggtopi_subs.genus)[rownames(res37.filt),])) %>%
+  cbind(as.matrix(tax_table(ggtopi_subs.genus)[rownames(res37.filt),-7])) %>%
   ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
@@
 ```
+
+### Boxplot of Genus abundances of differenitally abundant Genera
+```{r}
+  genus_ab_t <- otu_table(ggtopi_subs.genus.relab)[rownames(res37.filt), ] %>%
+              data.frame %>%
+              tibble::rownames_to_column("OTU") %>%
+              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
+              mutate(condition= if_else(timepoint=='T3','',stringr::str_sub(.$SampleID,4, 6)))
+
+  genus_ab_t %>% bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
+  ggplot() +
+    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=interaction(timepoint, condition))) + 
+    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
+    coord_flip() + 
+    theme_bw() +
+    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
+        strip.text.y = element_text(angle = 0))
+```

### 2018-11-26 cage_effect.Rmd
26112018

@@
 date: "November 8, 2018"
-output: html_document
+output: 
+  html_document: 
+    df_print: kable
+    theme: cerulean
 editor_options: 

### 2018-11-26 da_test.Rmd
26112018

@@
   html_document: 
-    number_sections: yes
     theme: cosmo
@@
   t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_filt), perl = TRUE)
-  tax_table(gg_topi.relab)[rownames(res37.filt),]
 
@@
 
-  genus_ab_t %>% bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
+  genus_ab_t %>%
+    bind_rows(genus_ab_t[genus_ab_t$timepoint=='T7',-5]) %>%
+    bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
     ggplot() +
@@
       theme_bw() +
-    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
+      theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
              strip.text.y = element_text(angle = 0))

### 2018-11-26 gg_topi.Rmd
26112018

@@
     df_print: kable
-    self_contained: no
-    theme: cerulean
+    theme: cosmo
 editor_options: 
-  chunk_output_type: inline
+  chunk_output_type: console
 ---
@@
 distancesjac$timepoint1 <- stringr::str_replace(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
+distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
 ```
@@
   geom_point(size = 0.8) +
-  geom_point(data=centroids, size = 4)
+  geom_point(data=centroids, size = 4) + 
+  theme_minimal()
 ```
@@
 
-## PERMANOVA all vs all
-```{r, cache=TRUE}
-permanova.res <- data.frame()
-for (s in combn(levels(sample_data(gg_topi)$cond_all),2, simplify = FALSE)) {
-  metadata_2cat <- data.frame(sample_data(gg_topi))[sample_data(gg_topi)$cond_all %in% s,]
-  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, cond_all %in% s), "wunifrac")
-  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat, parallel = 2)
-  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
-}
-```
-
-```{r}
-knitr::kable(permanova.res)
-```
-
 ## (T5+T6+T7) vs baseline
@@
 ```
+
 ```{r include=TRUE,  results = "asis"}
@@
 ```
+
+## PERMANOVA all vs all
+```{r, cache=TRUE}
+permanova.res <- data.frame()
+for (s in combn(levels(sample_data(gg_topi)$cond_all),2, simplify = FALSE)) {
+  metadata_2cat <- data.frame(sample_data(gg_topi))[sample_data(gg_topi)$cond_all %in% s,]
+  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, cond_all %in% s), "wunifrac")
+  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat, parallel = 2)
+  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
+}
+```
+
+```{r}
+knitr::kable(permanova.res)
+```
+
+# Abundance of Bacteroidales 24/4 through timepoints
+```{r}
+idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
+
+otu_table(gg_topi.genus.relab)[idx_247,] %>%
+  data.frame %>%
+  tibble::rownames_to_column("OTU") %>%
+  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
+  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
+  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
+  ggplot() +
+    geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
+    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
+    # coord_flip() + 
+    theme_bw() +
+    theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
+           strip.text.y = element_text(angle = 0))
+```
+
+# Correlation of tumor size at T7 and abundance of Bacteroidales 24/ through timepoints
+```{r}
+mass_247 <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
+  data.frame %>%
+  tibble::rownames_to_column("OTU") %>%
+  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
+  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
+  mutate(mouseID=as.integer(stringr::str_sub(.$SampleID,8, 10))) %>%
+  left_join(mappingfile[mappingfile$timepoint=='T7',], by=c('mouseID' = 'id-topo'))
+
+mass_247 %>%
+  ggplot(aes(Abundance,tumor_mass)) +
+  geom_point() + 
+  facet_grid(timepoint.x~condition, scales = "free", space = "free_y")+
+  geom_smooth(method = 'lm')
+```
+
+```{r}
+ds_corr_247 <- data.frame()
+for(x in unique(mass_247$timepoint.x)){
+  for(y in unique(mass_247$condition)){
+    tmp_res <- with(mass_247 %>% filter(timepoint.x==x & condition==y), cor.test(Abundance, tumor_mass))
+    ds_corr_247 %<>% bind_rows(data.frame(Cond = paste(y, ' ', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
+  }
+}
+print(knitr::kable(ds_corr_247))
+```
+
+# Correlation of tumor size at T7 and difference of abundance of Bacteroidales 24/7 through timepoints
+```{r}
+mass_247 %>%
+  filter(timepoint.x %in% c('T3','T7')) %>%
+  select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
+  tidyr::spread(timepoint.x, Abundance) %>%
+  mutate(T3_7 = abs(T7-T3)) %>%
+  ggplot(aes(T3_7,tumor_mass)) +
+  geom_point() +
+  geom_smooth(method = 'lm')
+```
+
+```{r}
+with(mass_247 %>%
+  filter(timepoint.x %in% c('T3','T7')) %>%
+  select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
+  tidyr::spread(timepoint.x, Abundance) %>%
+  mutate(T3_7 = abs(T7-T3)),
+  cor.test(T3_7,tumor_mass))
+
+## Ordination: T3 vs T7 with tumor size
+```{r}
+metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
+distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
+permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
+
+subset(ordwu, timepoint %in% c('T7','T3')) %>%
+ggplot(aes(Axis.1, Axis.2)) +
+  geom_point(aes(shape=timepoint, color=tumor_mass), size=3, alpha=.7, stroke=0.8) +
+  theme_minimal() +
+  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
+```

### 2019-01-06 paper_figures.Rmd
package setup. Figures.

---
title: "Paper figures"
author: "Francesco Beghini"
date: "January 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(phyloseq)
library(magrittr)
library(tidyverse)
library(reshape2)
library(ggsignif)
library(DESeq2)

source('R/loadQiimeData.R')
```

```{r QIIME data loading}
gg_topi <- loadQiimeData()
```

# Figure 3
## Panel A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3 
```{r Calculate alpha diversity}
alpha_div <- estimate_richness(gg_topi) %>% bind_cols(data.frame(sample_data(gg_topi)))
```

```{r}
filt_alpha <- alpha_div %>% 
  filter(timepoint %in% c('T1', 'T2', 'T3')) %>%
  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
  rename('se.chao1' = 'Chao1') %>%
  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2))) %>% droplevels()

ttest_res <- data.frame()
for ( ts in list(c('T1','T2'), c('T2','T3')) ){
  for( measure in c('Observed', 'Chao1', 'Simpson', 'Shannon') ){
    ttest_res %<>% bind_rows( data.frame(Measure = measure,
                                         Timepoint = paste(ts, collapse = ' vs '),
                                         pvalue = with(filt_alpha %>% filter(timepoint %in% ts), t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
  }
}
```

```{r}
knitr::kable(ttest_res)
```

```{r Plot boxplot}
filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'Chao1', 'Simpson', 'Shannon')) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, value)) +
  xlab('Timepoint') +
  ylab('Value') +
  facet_grid(variable~.,scales = 'free_y') +
  theme_minimal() +
  theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
           strip.text.y = element_text(angle = 0))
```

## Panel B: Beta diversity differences pre and post challange T2 vs T3
```{r}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
distanceswu <- calculate_pairwise_distances(distwu)
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T2','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T2','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
```

```{r Plot ordination}
subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))),
    gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = -.05, ymax = -.05)
```

## Panel C
```{r}

```

## Panel D
```{r}

```

### 2019-01-06 R/calculatePairwiseDistances.R

@@ -1,0 +1,12 @@
+calculate_pairwise_distances <- function(distmat){
+  distances <- reshape2::melt(as.matrix(distmat), varnames = c('rows','cols'))
+  distances$condizione1 <- sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][2])
+  distances$pentatome1 <- sapply(as.character(distances$rows), function(x) str_sub(x, 5,5))
+  distances$condizione2 <- sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][2])
+  distances$topo1 <- factor(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][3]))
+  distances$topo2 <- factor(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][3]))
+  distances$timepoint1 <- stringr::str_replace(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
+  distances$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
+
+  distances
+}

### 2019-01-06 R/loadQiimeData.R

@@ -1,0 +1,59 @@
+loadQiimeData <- function() {
+  gg_topi <- import_biom(BIOMfilename = 'inst/extdata/otu_table_mc10_w_tax_rdp.biom',
+                         treefilename = 'inst/extdata/rep_set.tre',
+                         refseqfilename = 'inst/extdata/rep_set.fna',
+                         refseqFunction = parse_taxonomy_default
+  )
+
+  mappingfile <- read.delim('inst/extdata/mappingfile', row.names = 1, check.names = FALSE)
+  mappingfile$Description %<>% str_replace_all(.,'_','-')
+  mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
+  sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
+  rownames(mappingfile) <- mappingfile$Description
+
+  sample_data(gg_topi) <- mappingfile
+  sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)
+
+  sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
+  sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])
+
+  sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
+  sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
+  sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))
+
+  sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)
+
+  colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
+
[…]

### 2019-01-06 da_test.Rmd
new figures. Fixed family levels boxplots

@@
 ```{r load qiime, include=FALSE, cache=TRUE}
-gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/otu_table_mc10_w_tax_rdp.biom',
-                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.tre',
-                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.fna',
-                        refseqFunction = parse_taxonomy_default
-                      )
-
-mappingfile <- read.delim('/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/mappingfile', row.names = 1, check.names = FALSE)
-mappingfile$Description %<>% str_replace_all(.,'_','-')
-mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
-sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
-rownames(mappingfile) <- mappingfile$Description
-
-sample_data(gg_topi) <- mappingfile
-sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)
-
-sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
-sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])
-
-sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
-sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
-sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))
-
-sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)
-
-colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
-
-tax_table_filtered <- tax_table(gg_topi) %>%
-  as.data.frame %>% 
-  mutate_all( .funs = function(x) stringr:::str_replace_all(x,"D_.__",""))
-
-rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
-tax_table(gg_topi) <-  as.matrix(tax_table_filtered)
-
-unass_otus <- tax_table(gg_topi) %>%
-                data.frame %>%
-                tibble::rownames_to_column('OTU') %>%
-                filter(Domain == 'Unassigned') %>%
-                select(OTU)
-
-unknown_otus <- tax_table(gg_topi) %>%
-                data.frame %>%
-                tibble::rownames_to_column('OTU') %>%
-                filter(is.na(Genus)) %>%
-                select(OTU)
-
-unknown_family <- tax_table(gg_topi) %>%
-                data.frame %>%
-                tibble::rownames_to_column('OTU') %>%
-                filter(is.na(Family)) %>%
-                select(OTU)
-
-tax_table(gg_topi)[unass_otus$OTU, ] <- rep.int('Unassigned', 7)
-tax_table(gg_topi)[unknown_otus$OTU, 'Genus'] <- 'Unknown'
-tax_table(gg_topi)[unknown_family$OTU, 'Family'] <- 'Unknown'
+gg_topi <- loadQiimeData()
 
@@
 
-gg_topi.family <- tax_glom(gg_topi,taxrank = "Genus")
+gg_topi.family <- tax_glom(gg_topi,taxrank = "Family")
 gg_topi.family.relab <- transform_sample_counts(gg_topi.family, function(x) x/sum(x))
@@
 ```{r}
-res.filt_family %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi.family)[rownames(res.filt_family),-7])) %>%
+res.filt_family %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi.family)[rownames(res.filt_family),c(-7,-6)])) %>%
   ggplot(aes(log2FoldChange, Family)) +

### 2019-01-06 gg_topi.Rmd
new figures. Fixed family levels boxplots

@@
 
+source('R/loadQiimeData.R')
+source('R/calculatePairwiseDistances.R')
 ```
@@
 ```{r load qiime, include=FALSE, cache=TRUE}
-gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/otu_table_mc10_w_tax_rdp.biom',
-                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.tre',
-                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.fna',
-                        refseqFunction = parse_taxonomy_default
-                      )
-
-mappingfile <- read.delim('/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/mappingfile', row.names = 1, check.names = FALSE)
-mappingfile$Description %<>% str_replace_all(.,'_','-')
-mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
-sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
-rownames(mappingfile) <- mappingfile$Description
-
-sample_data(gg_topi) <- mappingfile
-sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)
-
-sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
-sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])
-
-sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
-sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
-sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))
-
-sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)
-
-colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
-
-tax_table_filtered <- tax_table(gg_topi) %>%
-  as.data.frame %>% 
-  mutate_all( .funs = function(x) stringr:::str_replace_all(x,"D_.__",""))
-
-rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
-tax_table(gg_topi) <-  as.matrix(tax_table_filtered)
-
-unass_otus <- tax_table(gg_topi) %>%
-                data.frame %>%
-                tibble::rownames_to_column('OTU') %>%
-                filter(Domain == 'Unassigned') %>%
-                select(OTU)
-
-unknown_otus <- tax_table(gg_topi) %>%
-                data.frame %>%
-                tibble::rownames_to_column('OTU') %>%
-                filter(is.na(Genus)) %>%
-                select(OTU)
-
-tax_table(gg_topi)[unass_otus$OTU, ] <- rep.int('Unassigned', 7)
-tax_table(gg_topi)[unknown_otus$OTU, 'Genus'] <- 'Unknown'
+gg_topi <- loadQiimeData()
 
@@
 gg_topi.phylum.relab <- transform_sample_counts(gg_topi.phylum, function(x) x/sum(x))
-
-meta_samples <- data.frame(otu_table(gg_topi), check.names = FALSE) %>% 
-  tibble::rownames_to_column("OTU") %>%
-  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
-  tibble::add_column(., timepoint = stringr::str_sub(.$SampleID, 12, 13)) %>%
-  dplyr::group_by(timepoint, OTU) %>%
-  dplyr::select(-SampleID)%>%
-  dplyr::summarise(Abundance=mean(Abundance)) %>%
-  tidyr::spread(timepoint, Abundance) %>%
-  tibble::column_to_rownames(var = 'OTU') %>%
-  data.frame()
-
-gg_topi_meta_samples <- phyloseq(otu_table(meta_samples, taxa_are_rows = TRUE), tax_table(gg_topi), phy_tree(gg_topi))
 ```
@@
 
-# Alpha diversity
-```{r Alpha diversity}
-# otu_table(gg_topi) %>% colSums() %>% sort %>% barplot(xlab = c())
-# otu_table(gg_topi) %>% colSums() %>% summary()
-
+```{r Calculate alpha diversity}
 alpha_div <- estimate_richness(gg_topi)
 alpha_div %<>% bind_cols(data.frame(sample_data(gg_topi)))
+```
 
-x <- melt(alpha_div, id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon'))
-x$value <- as.numeric(x$value)
-
-ggplot(x) +
+# Alpha diversity: timepoint x Index x Condition
+```{r Plot Alpha diversity boxplots}
+melt(alpha_div, id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon')) %>%
+ggplot() +
   geom_boxplot(aes(timepoint, value)) +
@@
 ```
-<!-- ## SFB abundance -->
-<!-- ```{r} -->
-<!-- knitr::kable(otu_table(gg_topi)[c(879,880),]) -->
-<!-- ``` -->
+
+## Alpha diversity: tumor vs no tumor
+```{r}
+filt_alpha <- alpha_div %>% 
+  filter(timepoint %in% c('T1', 'T2', 'T3')) %>%
+  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
+  rename('se.chao1' = 'Chao1') %>%
+  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2))) %>% droplevels()
+
+ttest_res <- data.frame()
+for ( ts in list(c('T1','T2'), c('T2','T3')) ){
+  for( measure in c('Observed', 'Chao1', 'Simpson', 'Shannon') ){
+    ttest_res %<>% bind_rows( data.frame(Measure = measure,
+                                         Timepoint = paste(ts, collapse = ' vs '),
+                                         pvalue = with(filt_alpha %>% filter(timepoint %in% ts), t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
+  }
+}
+
+filt_alpha %>%
+  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'Chao1', 'Simpson', 'Shannon')) %>%
+  ggplot() +
+  geom_boxplot(aes(timepoint, value)) +
+  xlab('Timepoint') +
+  ylab('Value') +
+  facet_grid(variable~.,scales = 'free_y') +
+  theme_minimal() +
+  theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
+           strip.text.y = element_text(angle = 0))
+
+knitr::kable(ttest_res)
+```
 
 # Beta diversity
-```{r Beta diversity, cache=FALSE}
+```{r Beta diversity, cache=TRUE}
 distwu <- phyloseq::distance(gg_topi, "wunifrac")
@@
 distjac <- phyloseq::distance(gg_topi, "jaccard")
-distwu_meta <- phyloseq::distance(gg_topi_meta_samples, "jaccard")
 
 ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
-ordwu_meta <-  ordinate(gg_topi_meta_samples, method = "MDS", distance = distwu_meta, weighted = TRUE)
 ordbc <-  ordinate(gg_topi, method = "MDS", distance = distbc, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
@@
 
-# plot_ordination(gg_topi, ordwu, color = 'condizione',shape = 'timepoint', title = 'Weighted UniFrac')
-# plot_ordination(gg_topi, ordwu, color = 'merge_timepoint', title = 'Weighted UniFrac')
-
-distances <- reshape2::melt(as.matrix(distwu), varnames = c('rows','cols'))
-distances$condizione1 <- sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][2])
-distances$pentatome1 <- sapply(as.character(distances$rows), function(x) str_sub(x, 5,5))
-distances$condizione2 <- sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][2])
-distances$topo1 <- factor(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][3]))
-distances$topo2 <- factor(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][3]))
-distances$timepoint1 <- stringr::str_replace(sapply(as.character(distances$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-distances$timepoint2 <- stringr::str_replace(sapply(as.character(distances$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-
-distancesbc <- reshape2::melt(as.matrix(distbc), varnames = c('rows','cols'))
-distancesbc$condizione1 <- sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][2])
-distancesbc$condizione2 <- sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][2])
-distancesbc$topo1 <- factor(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][3]))
-distancesbc$topo2 <- factor(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][3]))
-distancesbc$timepoint1 <- stringr::str_replace(sapply(as.character(distancesbc$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-distancesbc$timepoint2 <- stringr::str_replace(sapply(as.character(distancesbc$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-
-distancesjac <- reshape2::melt(as.matrix(distjac), varnames = c('rows','cols'))
-distancesjac$condizione1 <- sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][2])
-distancesjac$condizione2 <- sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][2])
-distancesjac$topo1 <- factor(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][3]))
-distancesjac$topo2 <- factor(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][3]))
-distancesjac$timepoint1 <- stringr::str_replace(sapply(as.character(distancesjac$rows), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
-distancesjac$timepoint2 <- stringr::str_replace(sapply(as.character(distancesjac$cols), function(x) strsplit(x,'-')[[1]][4]), 'T', '') %>% as.numeric
+distanceswu <- calculate_pairwise_distances(distwu)
+distancesbc <- calculate_pairwise_distances(distbc)
+distancesjac <- calculate_pairwise_distances(distjac)
 ```
@@
 ggplot(aes(Axis.1, Axis.2)) +
-  geom_point(aes(color=tumor_injected), size=3, alpha=.8) +
-  scale_color_brewer(name = "Tumor injected", palette="Dark2") + 
+  geom_point(aes(color=timepoint), size=3, alpha=.8) +
+  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
   theme_minimal() +
-  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = -0.2, xmax = -.2, ymin = .25, ymax = .25)
-```
-
-## Alpha diversity: tumor vs no tumor
-```{r}
-filt_alpha <- alpha_div %>% 
-  filter(timepoint %in% c('T2','T3')) %>%
-  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
-  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2)))
-
-ttest_res <- data.frame()
-for(measure in c('Observed', 'se.chao1', 'Simpson', 'Shannon')){
-  ttest_res %<>% bind_rows(data.frame(Measure=measure, pvalue=with(filt_alpha, t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
-}
-
-filt_alpha %>%
-  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon')) %>%
-  ggplot() +
-  geom_boxplot(aes(timepoint, value)) +
-  facet_grid(variable~.,scales = 'free_y')
-
-knitr::kable(ttest_res)
+  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))),
+    gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = -.05, ymax = -.05)
 ```
@@
 ```{r Beta diversity between timepoints t1}
-distances_tmp <- distances
+distances_tmp <- distanceswu
 distances_tmp[distances_tmp$timepoint1 == 1 | distances_tmp$timepoint1 == 2, 'timepoint1'] <- 2
@@
 ```{r Beta diversity between timepoints tn-1}
-tn1 <- distances %>% 
+tn1 <- distanceswu %>% 
   filter(topo1==topo2) %>%
@@
 ```{r}
-t1 <- distances %>% 
+t1 <- distanceswu %>% 
   dplyr::filter(topo1==topo2) %>%
@@
 ```{r}
-t37 <- distances %>% 
+t37 <- distanceswu %>% 
   dplyr::filter(topo1==topo2) %>%
@@
 ```{r corr tumor wunifrac}
-distances %>% 
+distanceswu %>% 
   dplyr::filter(topo1==topo2) %>%
@@
 ```{r}
-wunif_ds <- distances %>% 
+wunif_ds <- distanceswu %>% 
   dplyr::filter(topo1==topo2) %>%
@@
 ```{r}
-dct <- distances %>%
+dct <- distanceswu %>%
   filter(timepoint1 == timepoint2) %>%
@@
 ```{r}
-dtc <- distances %>%
+dtc <- distanceswu %>%
   filter(condizione1 == condizione2) %>%

### 2019-01-06 paper_figures.Rmd
new figures. Fixed family levels boxplots

@@
 gg_topi <- loadQiimeData()
+gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
+gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))
+
+gg_topi.family <- tax_glom(gg_topi,taxrank = "Family")
+gg_topi.family.relab <- transform_sample_counts(gg_topi.family, function(x) x/sum(x))
 ```
@@
 
-## Panel C
+## Panel C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.
+```{r DESeq2 loading, cache=TRUE}
+pds_genus <- phyloseq_to_deseq2(gg_topi.genus, design = ~timepoint)
+dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE)
+res_genus <- DESeq2::results(dds_genus, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
+res.filt_genus <- res_genus[!is.na(res_genus$padj),]
+res.filt_genus <- res.filt_genus[res.filt_genus$padj<0.05,]
+
+t23_samples_idx_genus <- grep('T2|T3', sample_names(gg_topi.genus.relab), perl = TRUE)
+
+genus_ab_t <- otu_table(gg_topi.genus.relab)[rownames(res.filt_genus), t23_samples_idx_genus] %>%
+            data.frame %>%
+            tibble::rownames_to_column("OTU") %>%
+            tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+            mutate(timepoint=factor(stringr::str_sub(.$SampleID,12, 13)))
+```
+
 ```{r}
+lacno_bact_relabs <-genus_ab_t %>% bind_cols(data.frame(tax_table(gg_topi.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>% filter(Family %in% c('Bacteroidales S24-7 group') | Genus %in% c('Lachnospiraceae NK4A136 group'))
+
+lacno_bact_relabs %>% group_by(timepoint, interaction(Family, Genus)) %>% dplyr::summarise(Mean=mean(Abundance), SD = sd(Abundance)) %>% knitr::kable()
+
+lacno_bact_relabs %>% 
+  mutate(Taxonomy=interaction(Family, Genus)) %>%
+  select(Abundance, Taxonomy, timepoint) %>%
+  group_by(timepoint, Taxonomy) %>%
+  summarise(Abundance = list(Abundance)) %>% 
+  spread(timepoint, Abundance) %>% 
+  group_by(Taxonomy) %>%
+  dplyr::summarise("Wilcoxon p value"=wilcox.test(unlist(T2), unlist(T3), paired = TRUE)$p.value)  %>% knitr::kable()
+```
+
+```{r}
+  lacno_bact_relabs %>%
+  ggplot() +
+    geom_boxplot(aes(timepoint, Abundance, fill=timepoint)) + 
+    facet_wrap(interaction(Family, Genus) ~ ., scales = "free_x", as.table= T) +
+    theme_minimal() +
+    guides(fill=FALSE) +
+    theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
+           strip.text.x = element_text(size=12))
 

### 2019-01-06 paper_figures.Rmd
added panel E

@@
 dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE)
+```
+
+```{r DESeq2 T2 T3}
 res_genus <- DESeq2::results(dds_genus, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
@@
 
-## Panel D
+## Panel D: Ordination centroids beta diversity all timepoints 
 ```{r}
+x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
+
+centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)
+
+ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
+  geom_point(size = 0.8) +
+  geom_point(data=centroids, size = 4) + 
+  theme_minimal()
+```
+
+## Panel ?: Ordination beta diversity T3 vs T7
+```{r}
+metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
+distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
+permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
+
+subset(ordwu, timepoint %in% c('T7','T3')) %>%
+ggplot(aes(Axis.1, Axis.2)) +
+  geom_point(aes(color=timepoint), size=3, alpha=.8) +
+  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
+  theme_minimal() +
+  annotation_custom(grob = grid::textGrob(bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
+```
+
+```{r, cache=TRUE}
+ggtopi_subs <-subset_samples(gg_topi, timepoint %in% c('T3','T7'))
+ggtopi_subs.relab <- transform_sample_counts(ggtopi_subs, function(x) x/sum(x))
+ggtopi_subs.genus <- tax_glom(ggtopi_subs,taxrank = "Genus")
+ggtopi_subs.genus.relab <- transform_sample_counts(ggtopi_subs.genus, function(x) x/sum(x))
+
+mmtx <- model.matrix(object = ~timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, data = data.frame(ggtopi_subs@sam_data))
+idx <- which(apply(mmtx, 2, function(x) all(x==0)))
+mmtx <- mmtx[,-idx]
+
+pds_37 <- phyloseq_to_deseq2(ggtopi_subs, design = ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
+dds_37 <- DESeq2::DESeq(pds_37, parallel = TRUE, quiet = TRUE, full = mmtx)
+
+pds_genus <- phyloseq_to_deseq2(ggtopi_subs.genus, design =  ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
+dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE, full = mmtx)
+```
+## Panel E: Boxplot of abundance of significant genera from DESeq2 differential analysis
+```{r}
+res37 <- DESeq2::results(dds_genus, name = "timepointT7", alpha = 0.05)
+res37.filt <- res37[!is.na(res37$padj),]
+res37.filt <- res37.filt[res37.filt$padj<0.05,]
+
+t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_subs), perl = TRUE)
+
+genus_ab_t <- otu_table(ggtopi_subs.genus.relab)[rownames(res37.filt), ] %>%
+            data.frame %>%
+            tibble::rownames_to_column("OTU") %>%
+            tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+            mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))
+
+genus_ab_t %>%
+
+  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
+ggplot() +
+  geom_boxplot(aes(reorder(interaction(Family, Genus), Abundance, FUN = median), Abundance, fill=timepoint)) + 
+  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
+  coord_flip() + 
+  theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
+      strip.text.y = element_text(angle = 0))
 

### 2019-01-07 R/ggtheme.R
added theme function

@@ -1,0 +1,7 @@
+theme_cm <- function(){
+  theme_minimal() +
+  theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
+        axis.text = element_text(color = "black", size = 10),
+        strip.text.y = element_text(angle = 0, size = 10)) +
+  scale_fill_brewer(name = "Timepoint", palette="Dark2")
+}

### 2019-01-07 paper_figures.Rmd
added panel

@@
 output: html_document
+editor_options: 
+  chunk_output_type: console
 ---
@@
 source('R/loadQiimeData.R')
+source('R/ggtheme.R')
 ```
@@
 ```
-## Panel E: Boxplot of abundance of significant genera from DESeq2 differential analysis
+
+## Panel E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered
 ```{r}
@@
 genus_ab_t %>%
+  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
+  filter(OTU %in% (genus_ab_t %>%
+  group_by(OTU, timepoint) %>%
+  summarise(median=median(Abundance)) %>%
+  group_by(OTU) %>%
+  summarise(filt=min(median)>0.01) %>% filter(filt==TRUE))$OTU) %>%
+  filter(Phylum != "Unassigned") %>%
+  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
+ggplot() +
+  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint)) +
+  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
+  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
+  coord_flip() +
+  xlab('') +
+  ylab("Relative abundance") +
+  theme_cm()
+```
 
-  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
+## Panel F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment
+```{r}
+idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
+idx_lachno <- which(tax_table(gg_topi.genus.relab)[,'Genus'] == "Lachnospiraceae NK4A136 group" )
+
+otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
+  data.frame %>%
+  tibble::rownames_to_column("OTU") %>%
+  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
+  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
+  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
+  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
   ggplot() +
-  geom_boxplot(aes(reorder(interaction(Family, Genus), Abundance, FUN = median), Abundance, fill=timepoint)) + 
-  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
-  coord_flip() + 
-  theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
-      strip.text.y = element_text(angle = 0))
-
+  geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
+  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
+  facet_grid(taxlab ~ ., scales = "free", space = "free_x", as.table= T) +
+  theme_cm()
 ```

### 2019-01-07 R/ggtheme.R

@@ -1,7 +1,14 @@
+if(!requireNamespace("showtext", quietly = TRUE)){
+  install.packages('showtext', dependencies = TRUE)
+}
+library(showtext)
+font_add_google("Roboto", "Roboto")
+font_add_google("Montserrat", "Montserrat")
+
 theme_cm <- function(){
   theme_minimal() +
-  theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        axis.text = element_text(color = "black", size = 10),
-        strip.text.y = element_text(angle = 0, size = 10)) +
-  scale_fill_brewer(name = "Timepoint", palette="Dark2")
+  theme(text = element_text(family = "Montserrat", size = 10, colour = "black"),
+        axis.text = element_text(family = "Montserrat", size = 10, colour = "black"),
+        strip.background = element_rect(colour = "transparent", fill = "transparent"),
+        strip.text = element_text(angle = -1, family = "Montserrat", size = 10, colour = "black"))
 }

### 2019-01-07 paper_figures.Rmd
new panel

@@
 date: "January 6, 2019"
-output: html_document
+output: 
+  html_document: 
+    fig_caption: yes
 editor_options: 
@@
 source('R/ggtheme.R')
+source('R/calculatePairwiseDistances.R')
 ```
 
-```{r QIIME data loading}
+```{r QIIME data loading, cache=TRUE}
 gg_topi <- loadQiimeData()
+mappingfile <- data.frame(sample_data(gg_topi))
 gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
@@
 ## Panel A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3 
-```{r Calculate alpha diversity}
+```{r Calculate alpha diversity, cache=TRUE}
 alpha_div <- estimate_richness(gg_topi) %>% bind_cols(data.frame(sample_data(gg_topi)))
@@
 ## Panel B: Beta diversity differences pre and post challange T2 vs T3
-```{r}
+```{r, cache=TRUE}
 distwu <- phyloseq::distance(gg_topi, "wunifrac")
@@
 distanceswu <- calculate_pairwise_distances(distwu)
+distancesbc <- calculate_pairwise_distances(phyloseq::distance(gg_topi, "bray"))
 metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T2','T3'),]
@@
   theme_minimal() +
-  annotation_custom(grob = grid::textGrob(bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
+  annotation_custom(grob = grid::textGrob(bquote(atop(italic("p-value")==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
 ```
@@
   geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
-  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
-  facet_grid(taxlab ~ ., scales = "free", space = "free_x", as.table= T) +
-  theme_cm()
+  scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
+  facet_wrap(facets = 'taxlab', nrow = 2, scales = "free", shrink = FALSE) +
+  theme_cm() +
+  ylab('Relative abundance') +
+  xlab('Timepoint')
 ```
+
+## Panel G: Scatterplot of tumor size and difference of \\beta diversity 
+```{r}
+brayc_ds <- distancesbc %>% 
+  dplyr::filter(topo1==topo2) %>%
+  dplyr::filter(timepoint1==3) %>%
+  dplyr::filter(timepoint2>4) %>%
+  dplyr::select(cols, timepoint2, value) %>%
+  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))
+
+ds_corr_braycf <- data.frame()
+for(x in unique(brayc_ds$timepoint2)){
+  # for(y in levels(brayc_ds$condizione)){
+    tmp_res <- with(brayc_ds %>% filter(timepoint2==x), cor.test(value, tumor_mass))
+    ds_corr_braycf %<>% bind_rows(data.frame(timepoint = paste('T', x, '-T3',sep = ''), corr = signif(tmp_res$estimate,2), pval = signif(tmp_res$p.value,2), label = paste('p-value==',signif(tmp_res$p.value,2),'~italic(r)==', signif(tmp_res$estimate,2))))
+}
+```
+
+```{r corr tumor braycurtis}
+distancesbc %>% 
+  dplyr::filter(topo1==topo2) %>%
+  dplyr::filter(timepoint1==3) %>%
+  dplyr::filter(timepoint2>4) %>%
+  dplyr::select(cols, timepoint2, value) %>%
+  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
+  mutate(timepoint = paste('T',timepoint2,'-T3',sep = '')) %>%
+  ggplot(aes(value,tumor_mass)) +
+  geom_point() +
+  facet_wrap(facets = 'timepoint', ncol = 1) +
+  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
+  ylab(bquote('Tumor mass'~( mm^3))) +
+  xlab('Bray Curtis distance') +
+  theme_cm() +
+  geom_text(data = ds_corr_braycf, aes(x=0.4,y=2500, label=label), parse = TRUE, inherit.aes = FALSE)
+
+```

### 2019-01-07 R/ggtheme.R

@@ -10,5 +10,5 @@ theme_cm <- function(){
   theme(text = element_text(family = "Montserrat", size = 10, colour = "black"),
         axis.text = element_text(family = "Montserrat", size = 10, colour = "black"),
         strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text = element_text(angle = -1, family = "Montserrat", size = 10, colour = "black"))
+        strip.text = element_text(family = "Montserrat", size = 12, colour = "black"))
 }

### 2019-01-07 paper_figures.Rmd
Edit captions and img size

@@
 knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE)
-knitr::opts_chunk$set(fig.width=12, fig.height=8) 
+knitr::opts_chunk$set(fig.width=8, fig.height=6) 
 library(phyloseq)
@@
 # Figure 3
-## Panel A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3 
 ```{r Calculate alpha diversity, cache=TRUE}
@@
 
-```{r Plot boxplot}
+```{r Plot boxplot, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3'}
 filt_alpha %>%
@@
 
-## Panel B: Beta diversity differences pre and post challange T2 vs T3
 ```{r, cache=TRUE}
@@
 
-```{r Plot ordination}
+```{r Plot ordination, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3"}
 subset(ordwu, timepoint %in% c('T2','T3')) %>%
@@
   scale_color_brewer(name = "Timepoint", palette="Dark2") + 
-  theme_minimal() +
-  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))),
+  theme_cm() +
+  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))),
     gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = -.05, ymax = -.05)
@@
 
-## Panel C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.
 ```{r DESeq2 loading, cache=TRUE}
@@
 
-```{r}
+```{r fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families."}
   lacno_bact_relabs %>%
@@
 
-## Panel D: Ordination centroids beta diversity all timepoints 
-```{r}
+```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints"}
 x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
@@
 
-## Panel E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered
 ```{r}
@@
 
+```
+
+```{r fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered"}
 genus_ab_t %>%
@@
 
-## Panel F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment
-```{r}
+```{r fig.cap="Fig 3 F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment"}
 idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
@@
 
-## Panel G: Scatterplot of tumor size and difference of \\beta diversity 
 ```{r}
@@
 
-```{r corr tumor braycurtis}
+```{r corr tumor braycurtis, fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity "}
 distancesbc %>% 

### 2019-01-08 paper_figures.Rmd
panel

@@
 ttest_res <- data.frame()
-for ( ts in list(c('T1','T2'), c('T2','T3')) ){
+for ( ts in combn(levels(filt_alpha$timepoint), 2, simplify = FALSE)){
   for( measure in c('Observed', 'Chao1', 'Simpson', 'Shannon') ){
@@
                                          Timepoint = paste(ts, collapse = ' vs '),
-                                         pvalue = with(filt_alpha %>% filter(timepoint %in% ts), t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
+                                         pvalue = with(filt_alpha %>% filter(timepoint %in% ts), t.test(as.formula(paste(measure, 't', sep = '~')), alternative ='two.sided'))$p.value))
   }
@@
 
-```{r Plot boxplot, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3'}
+```{r fig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.height=5, fig.width=5}
+filt_alpha %>%
+  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Chao1')) %>%
+  ggplot(aes(timepoint, value, fill=timepoint)) +
+  geom_boxplot(width=0.8) +
+  xlab('Timepoint') +
+  ylab('Chao1 Index') +
+  scale_fill_manual(name = "Timepoint", values = cm_palette) +
+  ggsignif::geom_signif(comparisons = list(c('T1','T3')), y_position = 240, tip_length = 0, test = 't.test') +
+  ggsignif::geom_signif(comparisons = list(c('T2','T3')), y_position = 230, tip_length = 0, test = 't.test') +
+  guides(fill=FALSE) +
+  theme_cm()
+```
+
+```{r supplfig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3'}
 filt_alpha %>%
@@
   ggplot() +
-  geom_boxplot(aes(timepoint, value)) +
+  geom_boxplot(aes(timepoint, value, fill=timepoint)) +
+  scale_fill_manual(name = "Timepoint", values = cm_palette) +
   xlab('Timepoint') +
@@
   facet_grid(variable~.,scales = 'free_y') +
-  theme_minimal() +
-  theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
-           strip.text.y = element_text(angle = 0))
+  guides(fill=FALSE) +
+  theme_cm()
 ```
@@
   geom_point(aes(color=timepoint), size=3, alpha=.8) +
-  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
+  scale_color_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
   theme_cm() +
   annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))),
-    gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = -.05, ymax = -.05)
+    gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = .05, ymax = .05)
 ```
@@
     geom_boxplot(aes(timepoint, Abundance, fill=timepoint)) + 
+    scale_fill_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
     facet_wrap(interaction(Family, Genus) ~ ., scales = "free_x", as.table= T) +
-    theme_minimal() +
     guides(fill=FALSE) +
-    theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
-           strip.text.x = element_text(size=12))
-
+    theme_cm()
 ```
@@
 ```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints"}
-x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
+not12 <- gg_topi %>%
+  subset_samples(!(timepoint %in% c("T1", "T2")))
+
+x<-not12 %>%
+  ordinate(method = "MDS", distance =  phyloseq::distance(not12, "wunifrac"), weighted = TRUE) %>% plot_ordination(not12, ., justDF = TRUE)
 
 centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)
-
+centroids[1:4,c('X2','Y2')] <- centroids[-1,c(2:3)]
 ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
-  geom_point(size = 0.8) +
+  geom_point(size = 0.8, alpha=0.7) +
   geom_point(data=centroids, size = 4) + 
-  theme_minimal()
+  geom_segment(aes(x=Axis.1, y=Axis.2, xend=X2, yend=Y2), 
+               arrow = arrow(length = unit(0.1,'inches')),
+               data=centroids, group="1") + 
+  scale_color_manual(name = "Timepoint", values = cm_palette[-c(1,2)]) +
+  theme_cm()
 ```
@@
   geom_point(aes(color=timepoint), size=3, alpha=.8) +
-  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
+  scale_color_manual(name = "Timepoint", values = cm_palette[c(4,7)])+
   theme_minimal() +
@@
             tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
-            mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))
+            mutate(timepoint=factor(stringr::str_sub(.$SampleID,12, 13)))
 
@@
 
-```{r fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered"}
+```{r fig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered"}
 genus_ab_t %>%
@@
 ggplot() +
-  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint)) +
+  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85)) +
   facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
-  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
+  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
   coord_flip() +
@@
 
-```{r fig.cap="Fig 3 F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment"}
+```{r supplfig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis"}
+genus_ab_t %>%
+  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
+  filter(Phylum != "Unassigned") %>%
+  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
+ggplot() +
+  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85)) +
+  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
+  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
+  coord_flip() +
+  scale_y_log10() +
+  xlab('') +
+  ylab("Relative abundance") +
+  theme_cm() +
+  theme(strip.text.y = element_text(angle = 360),
+        legend.position = "bottom")
+```
+
+```{r fig3F, fig.cap="Fig 3 F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment"}
 idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
@@
   scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
+  ggsignif::geom_signif(comparisons = list(c('T2','T3')), y_position = 0.5, tip_length = 0, test = 't.test') +
   facet_wrap(facets = 'taxlab', nrow = 2, scales = "free", shrink = FALSE) +
@@
   xlab('Timepoint')
+```
+
+```{r}
+k<- otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
+   data.frame %>%
+   tibble::rownames_to_column("OTU") %>%
+   tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+   mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
+   mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
+   group_by(timepoint, condition, OTU) %>%
+   summarise(Abundance = list(Abundance)) %>%
+   spread(timepoint, Abundance)
+
+abs_test <- data.frame() 
+for(c in unique(k$condition)){
+  for(o in unique(k$OTU)){
+    for(t in combn(colnames(k)[-c(1:2)], 2, simplify = FALSE)){
+      l <-k %>%
+        filter(condition == c & OTU == o) %>%
+        select(t)
+      res <- t.test(unlist(l[,2]),unlist(l[,3]))
+      abs_test %<>% bind_rows( data.frame( Condition = c,
+                                           OTU = o,
+                                           Timepoint = paste(t, collapse = ' vs '),
+                                           p.value = signif(res$p.value,3),
+                                           Direction = sign(diff(res$estimate))
+                                         )
+                             )
+    }
+  }
+}
+```
+```{r}
+knitr::kable(abs_test)
 ```
@@
     tmp_res <- with(brayc_ds %>% filter(timepoint2==x), cor.test(value, tumor_mass))
-    ds_corr_braycf %<>% bind_rows(data.frame(timepoint = paste('T', x, '-T3',sep = ''), corr = signif(tmp_res$estimate,2), pval = signif(tmp_res$p.value,2), label = paste('p-value==',signif(tmp_res$p.value,2),'~italic(r)==', signif(tmp_res$estimate,2))))
+    ds_corr_braycf %<>% bind_rows(data.frame(timepoint = paste('T', x, '-T3',sep = ''), corr = signif(tmp_res$estimate,2), pval = signif(tmp_res$p.value,2), label = paste('p=',signif(tmp_res$p.value,2),'\nr=', signif(tmp_res$estimate,2), sep = '')))
 }
+ds_corr_braycf[,c('x','y')] <- list(c(.45,.45,.45),c(300,650,2000))
 ```
 
-```{r corr tumor braycurtis, fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity "}
+```{r fig3G, corr tumor braycurtis, fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity "}
 distancesbc %>% 
@@
   geom_point() +
-  facet_wrap(facets = 'timepoint', ncol = 1) +
+  facet_wrap(facets = 'timepoint', nrow = 1, shrink = TRUE, scales = 'free_y') +
   geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
@@
   theme_cm() +
-  geom_text(data = ds_corr_braycf, aes(x=0.4,y=2500, label=label), parse = TRUE, inherit.aes = FALSE)
+  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label), inherit.aes = FALSE)
+```
+
+```{r corr tumor braycurtis, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity "}
+
+otu_table(gg_topi.genus.relab)[idx_247,] %>%
+  data.frame %>%
+  tibble::rownames_to_column("OTU") %>%
+  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
+  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
+  filter(timepoint %in% c('T3','T7')) %>%
+  select(-SampleID, -OTU) %>%
+  group_by(condition) %>%
+  summarise(Abundance = list(Abundance)) %>%
+  spread(condition, Abundance)
+
+  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
+  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus)))
 

### 2019-01-08 R/ggtheme.R

@@ -12,3 +12,5 @@ theme_cm <- function(){
         strip.background = element_rect(colour = "transparent", fill = "transparent"),
         strip.text = element_text(family = "Montserrat", size = 12, colour = "black"))
 }
+
+cm_palette <- c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628')

### 2019-01-09 paper_figures.Rmd
Changed fig dimension

@@
 
-```{r supplfig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3'}
+```{r supplfig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.dim = c(5,5)}
 filt_alpha %>%
@@
 
-```{r Plot ordination, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3"}
+```{r Plot ordination, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(5,5)}
 subset(ordwu, timepoint %in% c('T2','T3')) %>%
@@
   annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))),
-    gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = .05, ymax = .05)
+    gp=grid::gpar(col="black", fontsize=11)),  xmin = 0.15, xmax = .15, ymin = .12, ymax = .12)
 ```
@@
 
-```{r fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families."}
+```{r fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(4,4)}
   lacno_bact_relabs %>%
@@
 
-```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints"}
+```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(5,6)}
 not12 <- gg_topi %>%
@@
 ## Panel ?: Ordination beta diversity T3 vs T7
-```{r}
+```{r, fig.dim = c(5, 6)}
 metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
@@
 
-```{r fig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered"}
+```{r fig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.height=8}
 genus_ab_t %>%
@@
 
-```{r supplfig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis"}
+```{r supplfig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis", fig.height=9}
 genus_ab_t %>%
@@
   scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
-  ggsignif::geom_signif(comparisons = list(c('T2','T3')), y_position = 0.5, tip_length = 0, test = 't.test') +
   facet_wrap(facets = 'taxlab', nrow = 2, scales = "free", shrink = FALSE) +
@@
 
-```{r fig3G, corr tumor braycurtis, fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity "}
+```{r fig3G, corr tumor braycurtis, fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
 distancesbc %>% 
@@
 
-```{r corr tumor braycurtis, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity "}
-
-otu_table(gg_topi.genus.relab)[idx_247,] %>%
+```{r corr tumor braycurtis, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
+tbl_data <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
   data.frame %>%
-  tibble::rownames_to_column("OTU") %>%
-  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+  tidyr::gather_("SampleID", "Abundance", colnames(.)) %>%
+  mutate(mouseid=stringr::str_sub(.$SampleID,8, 10)) %>%
   mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
@@
   filter(timepoint %in% c('T3','T7')) %>%
-  select(-SampleID, -OTU) %>%
+  select(-SampleID) %>%
+  spread(timepoint,Abundance) %>%
+  mutate(diff_abb = T3-T7) %>%
+  arrange(mouseid) %>%
+  bind_cols( gg_topi.genus.relab %>%
+              sample_data %>%
+              filter(timepoint == 'T7') %>%
+              select(id.topo, tumor_mass) %>%
+              arrange(id.topo) %>%
+              select(-id.topo))
+
+tbl_data %>%
+ggplot(aes(diff_abb, tumor_mass)) +
+  geom_point() +
+  facet_wrap(facets = 'condition', nrow = 1, scales = 'free') +
+  stat_smooth(method = 'lm', color='red') + 
+  theme_cm() +
+  xlab('Relative abundance (T3-T7)') +
+  ylab(bquote('Tumor mass'~( mm^3)))
+```
+
+```{r}
+tbl_data %>%
+  select(-T3,-T7, -mouseid) %>%
   group_by(condition) %>%
-  summarise(Abundance = list(Abundance)) %>%
-  spread(condition, Abundance)
-
-  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
-  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus)))
-
+  summarise(diff_abb = list(diff_abb), tumor_mass = list(tumor_mass)) %>%
+  group_by(condition) %>%
+  mutate(cor.test = paste(unlist(cor.test(unlist(diff_abb), unlist(tumor_mass))[c('p.value','estimate')]), collapse = ' ')) %>%
+  separate(cor.test, into = c('p.value','estimate'), sep = ' ')
 ```

### 2019-01-10 paper_figures.Rmd
figures done

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE)
+knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE, dev=c('svg', 'png'))
 knitr::opts_chunk$set(fig.width=8, fig.height=6) 
@@
     ttest_res %<>% bind_rows( data.frame(Measure = measure,
-                                         Timepoint = paste(ts, collapse = ' vs '),
-                                         pvalue = with(filt_alpha %>% filter(timepoint %in% ts), t.test(as.formula(paste(measure, 't', sep = '~')), alternative ='two.sided'))$p.value))
+                                         Timepoint1 = ts[1],
+                                         Timepoint2 = ts[2],
+                                         pvalue = with(filt_alpha %>% filter(timepoint %in% ts), t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
   }
@@
 ```{r fig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.height=5, fig.width=5}
+annot_df <- (ttest_res %>% filter(Measure == 'Chao1') %>% filter(pvalue < 0.05) %>% mutate(pvalue = paste('p=',signif(pvalue,2), sep = '')))
 filt_alpha %>%
   melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Chao1')) %>%
-  ggplot(aes(timepoint, value, fill=timepoint)) +
-  geom_boxplot(width=0.8) +
+  ggplot(aes(timepoint, value)) +
+  geom_boxplot(aes(fill=timepoint), width=0.8, alpha=.8) +
   xlab('Timepoint') +
@@
   scale_fill_manual(name = "Timepoint", values = cm_palette) +
-  ggsignif::geom_signif(comparisons = list(c('T1','T3')), y_position = 240, tip_length = 0, test = 't.test') +
-  ggsignif::geom_signif(comparisons = list(c('T2','T3')), y_position = 230, tip_length = 0, test = 't.test') +
+  # ggsignif::geom_signif(comparisons = list(c('T1','T3')), y_position = 240, tip_length = 0, test = 't.test') +
+  # ggsignif::geom_signif(comparisons = list(c('T2','T3')), y_position = 230, tip_length = 0, test = 't.test') +
+  geom_signif(data = annot_df, manual=TRUE, aes(annotations = pvalue, xmin=Timepoint1, xmax=Timepoint2, y_position = c(235,225))) +
   guides(fill=FALSE) +
@@
 
-```{r Plot ordination, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(5,5)}
+```{r Plot ordination, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(6,5)}
 subset(ordwu, timepoint %in% c('T2','T3')) %>%
@@
 ```{r}
-lacno_bact_relabs <-genus_ab_t %>% bind_cols(data.frame(tax_table(gg_topi.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>% filter(Family %in% c('Bacteroidales S24-7 group') | Genus %in% c('Lachnospiraceae NK4A136 group'))
+lacno_bact_relabs <- genus_ab_t %>% 
+                     bind_cols(data.frame(tax_table(gg_topi.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>% 
+                     filter(Family %in% c('Bacteroidales S24-7 group') | Genus %in% c('Lachnospiraceae NK4A136 group')) %>% 
+                     mutate(taxlab = str_replace(ifelse(stringr::str_detect(Genus,'uncultured'), paste(Family,sep = ''), paste(Genus, sep = '')), ' group','')) %>%
+                     select(taxlab, Abundance, timepoint, OTU)
 
-lacno_bact_relabs %>% group_by(timepoint, interaction(Family, Genus)) %>% dplyr::summarise(Mean=mean(Abundance), SD = sd(Abundance)) %>% knitr::kable()
+lacno_bact_relabs %>% group_by(timepoint,taxlab) %>% dplyr::summarise(Mean=mean(Abundance), SD = sd(Abundance)) %>% knitr::kable()
 
-lacno_bact_relabs %>% 
-  mutate(Taxonomy=interaction(Family, Genus)) %>%
-  select(Abundance, Taxonomy, timepoint) %>%
-  group_by(timepoint, Taxonomy) %>%
-  summarise(Abundance = list(Abundance)) %>% 
-  spread(timepoint, Abundance) %>% 
-  group_by(Taxonomy) %>%
-  dplyr::summarise("Wilcoxon p value"=wilcox.test(unlist(T2), unlist(T3), paired = TRUE)$p.value)  %>% knitr::kable()
 ```
 
-```{r fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(4,4)}
-  lacno_bact_relabs %>%
-  ggplot() +
-    geom_boxplot(aes(timepoint, Abundance, fill=timepoint)) + 
-    scale_fill_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
-    facet_wrap(interaction(Family, Genus) ~ ., scales = "free_x", as.table= T) +
-    guides(fill=FALSE) +
-    theme_cm()
+```{r fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(6,5)}
+annot_data <- res.filt_genus[unique(lacno_bact_relabs$OTU),] %>% 
+              data.frame %>% 
+              rownames_to_column('OTU') %>% 
+              bind_cols(taxlab = c('Lachnospiraceae NK4A136','Bacteroidales S24-7'),  xmin = c('T2','T2'), timepoint = c('T3','T3'), y_position =  c(0.6,0.7)) %>%
+              select(annotations = padj, taxlab, timepoint, xmin, y_position)
+
+ggplot(data=lacno_bact_relabs, aes(x=timepoint, y=Abundance)) +
+  stat_boxplot(aes(fill=timepoint), alpha=.8) +
+  facet_wrap(~taxlab) +
+  scale_fill_manual('Timepoint', values=cm_palette[c(2,3)]) + 
+  geom_signif(data = annot_data,
+              aes(annotations = signif(annotations,2), xmin=xmin, xmax=timepoint, y_position = y_position), manual = TRUE) +
+  theme_cm() +
+  theme(strip.text = element_text(size=11)) +
+  xlab('Timepoint') +
+  ylab('Relative abundance')+
+  guides(fill=FALSE)
 ```
 
-```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(5,6)}
+```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(6,5)}
 not12 <- gg_topi %>%
@@
 ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
-  geom_point(size = 0.8, alpha=0.7) +
+  geom_point(size = 1.5, alpha=0.7) +
   geom_point(data=centroids, size = 4) + 
@@
 ## Panel ?: Ordination beta diversity T3 vs T7
-```{r, fig.dim = c(5, 6)}
+```{r, fig.dim = c(6, 5)}
 metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
@@
   geom_point(aes(color=timepoint), size=3, alpha=.8) +
-  scale_color_manual(name = "Timepoint", values = cm_palette[c(4,7)])+
+  scale_color_manual(name = "Timepoint", values = cm_palette[c(3,7)])+
   theme_minimal() +
-  annotation_custom(grob = grid::textGrob(bquote(atop(italic("p-value")==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
+  annotation_custom(grob = grid::textGrob(bquote(atop(italic("p-value")==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=11)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
 ```
@@
 
-```{r fig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.height=8}
+```{r fig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.height=6}
 genus_ab_t %>%
@@
 ggplot() +
-  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85)) +
+  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85),alpha=.8) +
   facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
@@
 
+```{r}
+idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
+idx_lachno <- which(tax_table(gg_topi.genus.relab)[,'Genus'] == "Lachnospiraceae NK4A136 group" )
+
+k<- otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
+   data.frame %>%
+   tibble::rownames_to_column("OTU") %>%
+   tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
+   mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
+   mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
+   group_by(timepoint, condition, OTU) %>%
+   summarise(Abundance = list(Abundance)) %>%
+   spread(timepoint, Abundance) %>%
+   bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Family','Genus')])) %>%
+   mutate(taxlab = str_replace(ifelse(stringr::str_detect(Genus,'uncultured'), paste(Family,sep = ''), paste(Genus, sep = '')), ' group','')) %>%
+   select(-Family, -Genus)
+
+abs_test <- data.frame() 
+for(c in unique(k$condition)){
+  for(o in unique(k$taxlab)){
+    for(t in combn(colnames(k)[-c(1:2,10)], 2, simplify = FALSE)){
+      l <-k %>%
+        filter(condition == c & taxlab == o) %>%
+        select(t)
+      res <- t.test(unlist(l[,2]),unlist(l[,3]))
+      abs_test %<>% bind_rows( data.frame( Condition = c,
+                                           taxlab = o,
+                                           Timepoint1 = t[1],
+                                           Timepoint2 = t[2],
+                                           p.value = signif(res$p.value,3),
+                                           Direction = sign(diff(res$estimate))
+                                         )
+                             )
+    }
+  }
+}
+```
+
+```{r}
+knitr::kable(abs_test)
+```
+
 ```{r fig3F, fig.cap="Fig 3 F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment"}
@@
 idx_lachno <- which(tax_table(gg_topi.genus.relab)[,'Genus'] == "Lachnospiraceae NK4A136 group" )
+
+annot_df <- abs_test %>%
+  filter(Condition %in% c('VVB','VVP')) %>%
+  filter(Timepoint2 == 'T6' & Timepoint1 %in% c('T4','T5')) %>%
+  mutate(xmin = as.numeric(str_sub(Timepoint1, 2,2)) + ifelse(Condition == 'VVB', .094, .282),
+         xmax = as.numeric(str_sub(Timepoint2, 2,2)) + ifelse(Condition == 'VVB', .094, .282),
+                    #L4VVB,L5VVB,B4VVB,B5VVB,L4VVP,L5VVP,B4VVP,B5VVP
+         y_position = c(.75,  #L4B
+                        .45, #L5B
+
+                        .93,  #B4B
+                        .73,  #B5B
+
+                        .65,  #L4P
+                        .55,  #L5P
+
+                        .83, #B4P
+                        .63) #B5P
+         ,
+         annotations = paste('p=',signif(p.value, 2), sep = '')) %>%
+  select(taxlab, annotations, xmax, xmin, y_position)
 
@@
   bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
-  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
+  mutate(taxlab = str_replace(ifelse(stringr::str_detect(Genus,'uncultured'), paste(Family,sep = ''), paste(Genus, sep = '')), ' group','')) %>%
 ggplot() +
-  geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
+  geom_boxplot(aes(timepoint, Abundance, fill=condition), alpha=.8) +
   scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
-  facet_wrap(facets = 'taxlab', nrow = 2, scales = "free", shrink = FALSE) +
+  facet_wrap(facets = 'taxlab', nrow = 2, scales = "free_x") +
+  geom_signif(data = annot_df, manual = TRUE, aes(xmin = xmin, xmax = xmax, annotations = annotations, y_position = y_position),textsize=3) +
   theme_cm() +
@@
   xlab('Timepoint')
-```
-
-```{r}
-k<- otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
-   data.frame %>%
-   tibble::rownames_to_column("OTU") %>%
-   tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
-   mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
-   mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
-   group_by(timepoint, condition, OTU) %>%
-   summarise(Abundance = list(Abundance)) %>%
-   spread(timepoint, Abundance)
-
-abs_test <- data.frame() 
-for(c in unique(k$condition)){
-  for(o in unique(k$OTU)){
-    for(t in combn(colnames(k)[-c(1:2)], 2, simplify = FALSE)){
-      l <-k %>%
-        filter(condition == c & OTU == o) %>%
-        select(t)
-      res <- t.test(unlist(l[,2]),unlist(l[,3]))
-      abs_test %<>% bind_rows( data.frame( Condition = c,
-                                           OTU = o,
-                                           Timepoint = paste(t, collapse = ' vs '),
-                                           p.value = signif(res$p.value,3),
-                                           Direction = sign(diff(res$estimate))
-                                         )
-                             )
-    }
-  }
-}
-```
-```{r}
-knitr::kable(abs_test)
 ```
@@
 }
-ds_corr_braycf[,c('x','y')] <- list(c(.45,.45,.45),c(300,650,2000))
+ds_corr_braycf[,c('x','y')] <- list(c(.45,.45,.45),c(300,670,2000))
 ```
@@
   facet_wrap(facets = 'condition', nrow = 1, scales = 'free') +
-  stat_smooth(method = 'lm', color='red') + 
+  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
   theme_cm() +

### 2019-01-10 R/ggtheme.R

@@ -13,4 +13,4 @@ theme_cm <- function(){
         strip.text = element_text(family = "Montserrat", size = 12, colour = "black"))
 }
 
-cm_palette <- c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#ffff33','#a65628')
+cm_palette <- c('#e41a1c','#377eb8','#4daf4a','#984ea3','#FF8F33','#FFE133','#6D3100')

### 2019-01-18 paper_figures.Rmd
edits

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE, dev=c('svg', 'png'))
+knitr::opts_chunk$set(
+  message = FALSE, 
+  warning = FALSE, 
+  paged.print=FALSE, 
+  echo=FALSE, 
+  fig.path = 'plots/',
+  dev=c('svg', 'png'))
 knitr::opts_chunk$set(fig.width=8, fig.height=6) 
@@
 
-```{r}
+```{r, cache=TRUE}
 filt_alpha <- alpha_div %>% 
@@
 
-```{r}
+```{r, cache=TRUE}
 knitr::kable(ttest_res)
@@
 
-```{r fig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.height=5, fig.width=5}
+```{r fig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.height=3, fig.width=3}
 annot_df <- (ttest_res %>% filter(Measure == 'Chao1') %>% filter(pvalue < 0.05) %>% mutate(pvalue = paste('p=',signif(pvalue,2), sep = '')))
@@
 
-```{r Plot ordination, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(6,5)}
+```{r fig3b, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(6,5)}
 subset(ordwu, timepoint %in% c('T2','T3')) %>%
@@
 
-```{r DESeq2 T2 T3}
+```{r DESeq2 T2 T3, cache=TRUE}
 res_genus <- DESeq2::results(dds_genus, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
@@
 
-```{r}
+```{r, cache=TRUE}
 lacno_bact_relabs <- genus_ab_t %>% 
@@
 
-```{r fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(6,5)}
+```{r fig3c, fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(6,5)}
 annot_data <- res.filt_genus[unique(lacno_bact_relabs$OTU),] %>% 
@@
 
-```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(6,5)}
+```{r fig3d, fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(6,5)}
 not12 <- gg_topi %>%
@@
 ## Panel ?: Ordination beta diversity T3 vs T7
-```{r, fig.dim = c(6, 5)}
+```{r fig3x, fig.dim = c(6, 5)}
 metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
@@
 
-mmtx <- model.matrix(object = ~timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, data = data.frame(ggtopi_subs@sam_data))
+mmtx <- model.matrix(object = ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, data = data.frame(ggtopi_subs@sam_data))
 idx <- which(apply(mmtx, 2, function(x) all(x==0)))
@@
 
-```{r}
+```{r, cache=TRUE}
 res37 <- DESeq2::results(dds_genus, name = "timepointT7", alpha = 0.05)
@@
 
-```{r fig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.height=6}
+```{r fig3e, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.height=6}
 genus_ab_t %>%
@@
   mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
+  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown'), paste('Unknown',Family,sep = ' '), paste(taxlab))) %>%
+  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown Unknown'), paste('Unknown',Phylum,sep = ' '), paste(taxlab))) %>%
+  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Uncultured uncultured'), paste('Uncultured',Phylum,sep = ' '), paste(taxlab))) %>%
 ggplot() +
@@
 
-```{r supplfig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis", fig.height=9}
+```{r supplfig3e, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis", fig.height=9}
 genus_ab_t %>%
@@
 
-```{r}
+```{r, cache=TRUE}
 idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
@@
 
-```{r fig3F, fig.cap="Fig 3 F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment"}
+```{r fig3f, fig.cap="Fig 3 F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment"}
 idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
@@
 
-```{r}
+```{r, cache=TRUE}
 brayc_ds <- distancesbc %>% 
@@
 
-```{r fig3G, corr tumor braycurtis, fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
+```{r fig3g, include=FALSE, eval=FALSE, fig.height=3,fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity "}
 distancesbc %>% 
@@
   theme_cm() +
-  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label), inherit.aes = FALSE)
+  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label))
 ```
 
-```{r corr tumor braycurtis, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
+```{r fig3i, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
 tbl_data <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
@@
 
-```{r}
+```{r, cache=TRUE}
 tbl_data %>%

### 2019-01-18 R/ggtheme.R

@@ -1,16 +1,17 @@
-if(!requireNamespace("showtext", quietly = TRUE)){
-  install.packages('showtext', dependencies = TRUE)
-}
-library(showtext)
-font_add_google("Roboto", "Roboto")
-font_add_google("Montserrat", "Montserrat")
+# if(!requireNamespace("showtext", quietly = TRUE)){
+#   install.packages('showtext', dependencies = TRUE)
+#   install.packages("extrafont")
+# }
+# library(showtext)
+# font_add_google("Roboto", "Roboto")
+# font_add_google("Montserrat", "Montserrat")
 
 theme_cm <- function(){
   theme_minimal() +
-  theme(text = element_text(family = "Montserrat", size = 10, colour = "black"),
-        axis.text = element_text(family = "Montserrat", size = 10, colour = "black"),
+  theme(text = element_text(family = "sans-serif", size = 10, colour = "black"),
+        axis.text = element_text(family = "sans-serif", size = 10, colour = "black"),
         strip.background = element_rect(colour = "transparent", fill = "transparent"),
-        strip.text = element_text(family = "Montserrat", size = 12, colour = "black"))
+        strip.text = element_text(family = "sans-serif", size = 12, colour = "black"))
 }
 
 cm_palette <- c('#e41a1c','#377eb8','#4daf4a','#984ea3','#FF8F33','#FFE133','#6D3100')

### 2019-02-14 gg_topi.Rmd
added panel
Added figures

@@
 gg_topi <- loadQiimeData()
-
+mappingfile <- data.frame(sample_data(gg_topi))
 gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
@@
   mutate(mouseID=as.integer(stringr::str_sub(.$SampleID,8, 10))) %>%
-  left_join(mappingfile[mappingfile$timepoint=='T7',], by=c('mouseID' = 'id-topo'))
+  mutate(mouseID=as.factor(mouseID))%>%
+  left_join(mappingfile[mappingfile$timepoint=='T7',], by=c('mouseID' = 'id.topo'))
 
@@
 mass_247 %>%
-  filter(timepoint.x %in% c('T3','T7')) %>%
+  filter(timepoint.x %in% c('T3','T5','T6','T7')) %>% 
   select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
   tidyr::spread(timepoint.x, Abundance) %>%
-  mutate(T3_7 = abs(T7-T3)) %>%
-  ggplot(aes(T3_7,tumor_mass)) +
+  mutate(T3_7 = abs(T7-T3), T3_6 = abs(T6-T3),T3_5 = abs(T5-T3)) %>%
+  select(-T3,-T5,-T6,-T7) %>%
+  tidyr::gather(key = timepoint, value=abundance, -condition, -tumor_mass, -mouseID) %>%
+  ggplot(aes(abundance, tumor_mass)) +
   geom_point() +
-  geom_smooth(method = 'lm')
+  facet_grid(timepoint~condition)+
+  geom_smooth(method = 'lm') + 
+  theme_cm() +
+
 ```
@@
   cor.test(T3_7,tumor_mass))
+
+```
 

### 2019-02-14 paper_figures.Rmd
Added figures

@@
 
-```{r fig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.height=3, fig.width=3}
+```{r fig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.height=3, fig.width=2}
 annot_df <- (ttest_res %>% filter(Measure == 'Chao1') %>% filter(pvalue < 0.05) %>% mutate(pvalue = paste('p=',signif(pvalue,2), sep = '')))
@@
   ggplot(aes(timepoint, value)) +
-  geom_boxplot(aes(fill=timepoint), width=0.8, alpha=.8) +
+  geom_boxplot(aes(fill=timepoint), width=0.5, alpha=.8) +
   xlab('Timepoint') +
@@
 
-```{r fig3b, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(6,5)}
+```{r fig3b, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(5,3)}
 subset(ordwu, timepoint %in% c('T2','T3')) %>%
@@
 
-```{r fig3c, fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(6,5)}
+```{r fig3c, fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(3,3)}
 annot_data <- res.filt_genus[unique(lacno_bact_relabs$OTU),] %>% 
@@
 ggplot(data=lacno_bact_relabs, aes(x=timepoint, y=Abundance)) +
-  stat_boxplot(aes(fill=timepoint), alpha=.8) +
+  stat_boxplot(aes(fill=timepoint), alpha=.8, width=0.5) +
   facet_wrap(~taxlab) +
@@
 
-```{r fig3d, fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(6,5)}
+```{r, cache=TRUE}
 not12 <- gg_topi %>%
@@
 centroids[1:4,c('X2','Y2')] <- centroids[-1,c(2:3)]
+```
+
+```{r fig3d, fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(5,3)}
 ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
@@
 ## Panel ?: Ordination beta diversity T3 vs T7
-```{r fig3x, fig.dim = c(6, 5)}
+```{r fig3x, fig.dim = c(5, 3)}
 metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
@@
   scale_color_manual(name = "Timepoint", values = cm_palette[c(3,7)])+
-  theme_minimal() +
+  theme_cm() +
   annotation_custom(grob = grid::textGrob(bquote(atop(italic("p-value")==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=11)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
@@
 
-```{r fig3e, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.height=6}
+```{r fig4a, fig.cap="Fig 4A: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.dim = c(5,3)}
 genus_ab_t %>%
@@
   filter(Phylum != "Unassigned") %>%
-  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
+  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('',Family,sep = ''), paste(Genus))) %>%
   mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown'), paste('Unknown',Family,sep = ' '), paste(taxlab))) %>%
@@
   mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Uncultured uncultured'), paste('Uncultured',Phylum,sep = ' '), paste(taxlab))) %>%
+  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'group'), str_replace(taxlab, ' group',''), paste(taxlab))) %>%
 ggplot() +
@@
 
-```{r supplfig3e, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis", fig.height=9}
+```{r supplfig4a, fig.cap="Supplemenrary Fig 4a: Boxplot of abundance of significant genera from DESeq2 differential analysis", fig.height=9}
 genus_ab_t %>%
@@
 
-```{r fig3f, fig.cap="Fig 3 F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment"}
+```{r fig4B, fig.cap="Fig 4B: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment", fig.dim=c(5,4)}
 idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
@@
   theme_cm() +
+  theme(strip.text = element_text(family = "sans-serif", size = 9, colour = "black"),
+        axis.text = element_text(family = "sans-serif", size = 8, colour = "black"),
+        axis.title = element_text(family = "sans-serif", size = 8, colour = "black"))+
   ylab('Relative abundance') +
@@
 
-```{r fig3g, include=FALSE, eval=FALSE, fig.height=3,fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity "}
+```{r fig4C, fig.cap="Fig 4C: Scatterplot of tumor size and difference of beta diversity ", fig.dim=c(5,2)}
 distancesbc %>% 
@@
 ggplot(aes(value,tumor_mass)) +
-  geom_point() +
+  geom_point(size=2) +
   facet_wrap(facets = 'timepoint', nrow = 1, shrink = TRUE, scales = 'free_y') +
   geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
-  ylab(bquote('Tumor mass'~( mm^3))) +
+  ylab('Tumor mass (mm3)') +
   xlab('Bray Curtis distance') +
+  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label), size=2) +
   theme_cm() +
-  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label))
+  theme(strip.text = element_text(family = "sans-serif", size = 9, colour = "black"),
+        axis.text = element_text(family = "sans-serif", size = 8, colour = "black"),
+        axis.title = element_text(family = "sans-serif", size = 8, colour = "black"))
 ```
 
-```{r fig3i, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
+```{r}
 tbl_data <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
@@
               select(-id.topo))
-
+```
+```{r fig4D, fig.cap="Fig 4D: Scatterplot of tumor size and difference of beta diversity ", fig.dim=c(3.5,3.5)}
 tbl_data %>%
 ggplot(aes(diff_abb, tumor_mass)) +
-  geom_point() +
-  facet_wrap(facets = 'condition', nrow = 1, scales = 'free') +
+  geom_point(size=2) +
+  facet_wrap(facets = 'condition', nrow = 2, scales = 'free') +
   geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
   theme_cm() +
+  theme(strip.text = element_text(family = "sans-serif", size = 9, colour = "black"),
+        axis.text = element_text(family = "sans-serif", size = 8, colour = "black"),
+        axis.title = element_text(family = "sans-serif", size = 8, colour = "black")) +
   xlab('Relative abundance (T3-T7)') +
-  ylab(bquote('Tumor mass'~( mm^3)))
+  ylab('Tumor mass (mm3)')
 ```
